---
title: "2019 IYS - Salmon Diet Data wrangle"
author: "Tim van der Stap"
date: "7/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(janitor)
library(readxl)
library(here)
library(parsedate)
library(googledrive)
library(uuid)
```

The raw data is stored on Google Drive. The first step is to download the file from Google Drive to a local folder on your computer's hard drive. 

``` {r file_download, eval = FALSE}
# Make sure your folder path exists already (e.g. ./Salmon Diet/raw_data)
drive_download("https://docs.google.com/spreadsheets/d/1E88ZXuiYIqjq4ABl2GDB7Kp0UAG-0tIM/edit#gid=2020388317", 
               path = here::here("Salmon Diet", "raw_data", 
                                 "GoA_fish diets_cleaned.xlsx"),
               overwrite = TRUE)
```

*** 

First step, is we wanted to see if the data in the salmon diet dataset matches data recorded in the pathogen dataset. The reason for this is that in the pathogen dataset, Floytags and MGL# are attached to the samples, but this appears to be missing in the salmon diet dataset. By cross-referencing, we can gain insight as to which diet data belongs to which Floytag or MGL#. Therefore, we also download the pathogen data.

``` {r file_download2, eval = FALSE}
# Make sure your folder path exists already (e.g. /Salmon Diet/raw_data)
drive_download("https://docs.google.com/spreadsheets/d/14YOHGGrU0zi--QHurgC6YVqlJ2GMPdoa/edit#gid=1919082779",
               path = here::here("Salmon Diet", "raw_data",
                                 "IYS_GoA_2019_fish_health_pathogen_detections.xlsx"),
               overwrite = TRUE)
```

Now read both files from the local folder, and make the minor adjustments in the data (column headers) needed to cross-reference columns `Trawl`, `Species` and `Mass`. 

``` {r, eval = FALSE}
salmon <- read_excel(here("Salmon Diet", "raw_data", 
                          "GoA_fish diets_cleaned.xlsx"),
                     sheet = "Diets_clean") %>%
  rename(Mass = `Mass, g`) %>%
  select(Trawl, Species, Mass)

pathogen <- read_excel(here("Salmon Diet", "raw_data",
                            "IYS_GoA_2019_fish_health_pathogen_detections.xlsx")) %>%
  mutate(Species = case_when(
    Species == "Coho" ~ "Oncorhynchus kisutch",
    Species == "Pink" ~ "Oncorhynchus gorbuscha",
    Species == "Chum" ~ "Oncorhynchus keta",
    Species == "Chinook" ~ "Oncorhynchus tschawytschka",
    Species == "Sockeye" ~ "Oncorhynchus nerka"
  )) %>%
  rename(Trawl = Set) %>%
  select(Trawl, Species, Mass)

# The rows where the is a match can be found using: 
match <- inner_join(salmon, pathogen) %>% mutate(check = "match")
# The rows where there is no match can be found using:
nomatch <- anti_join(salmon, pathogen) %>% mutate(check = "nomatch")
full <- rbind(match, nomatch)
full <- full[order(full$Trawl),]
```

Using the nrow() function we find that there are 52 rows where `Species`, `Trawl` and `Mass` data matches between the salmon diet data and the pathogen data, but 96 rows where there is no match. Unfortunately this means that, in the current state, we cannot cross-reference the pathogen and salmon data for the majority of the fish, and consequently we are not able to use Floytag or MGL# as a common parentEvent. 

***

# Event Core

When creating the different eventID layers, we assume that the column `Trawl` is synonymous to `Station`. Ideally, the salmon that were caught and dissected would have been given a Floytag or MGL#, so a distinction can be made between stomach content of different individuals of a same species in the same trawl. 

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 10px;}
</style>
<div class = "blue">

Current problem: Multiple individuals of the same species were caught in a single trawl, however within the salmon diet dataset there is no column to indicate how these individual measurements are separate (i.e., no sampleID associated to each individual fish). Additionally, some species' stomach content seems to have been aggregated, whereas others haven't. Even within a single trawl record there is data specific to a single individual and aggregated data, _and sometimes of the same species_. 

</div>

**Food for thought**: Currently I have the "Salmon caught" be an `event` layer, rather than an `Occurrence`. The measurements associated to the fish, such as `fork length`, `mass` etc are directly linked to the salmon individual. The reason why I currently included the salmon individual in the `Event` Core is because I believe the focus of the dataset is _prey occurrence and prey mass_ in the stomach content of salmon, and to examine whether viewing the salmon as a _sampling instrument_ is a viable option. 

``` {r data wrangle, eval = FALSE}
salmon <- read_excel(here("Salmon Diet", "raw_data", 
                       "GoA_fish diets_cleaned.xlsx"), 
                  sheet = "Diets_clean") %>%
  janitor::clean_names() %>%
  mutate(cruise = "GoA2019",
         station = paste(cruise, trawl, sep="_Stn"),
         trawl = paste(station, "trawl1", sep=":"),
         sample = paste(trawl, species, sep=":")
  )
```

No metadata is mentioned in the Salmon Diet data as to when and where the trawls took place. To connect the coordinates and time to the eventIDs, this information is taken from the trawl datasheet. Right now I am saving both these files in the Salmon Diet folder, but once all folders and scripts have been QC'd the location/time can be read straight from the Trawl-specific folder. For now, I'm afraid it would cause confusion when committing/pushing to GitHub. 

``` {r salmon_datetime, eval = FALSE}
drive_download(file = "https://docs.google.com/spreadsheets/d/1rtT6L5WSi_UiiXfoS95wnDqA8loc959T/edit#gid=668595086", 
               path = here::here("Salmon Diet", "raw_data", 
                                 "2019_GoA_Fish_data_Hakai.xlsx"),
               overwrite = TRUE)

trawl <- read_excel(here("Salmon Diet", "raw_data", "2019_GoA_Fish_data_Hakai.xlsx")) %>%
  mutate(eventDate = format_iso_8601(as.POSIXct(UTC_start)),
         eventDate = str_replace(eventDate, "\\+00:00", "Z")) %>%
  mutate(station = paste("GoA2019", NUMBER, sep = "_Stn"))
```

**Please note that the date and coordinates could change as the trawl data is not finalized.** 

``` {r salmondiet_event, eval = FALSE}
salmon_cruise <- salmon %>% 
  select(eventID = cruise) %>%
  distinct(eventID) %>%
  mutate(eventRemarks = "cruise")

salmon_station <- salmon %>%
  select(eventID = station,
         parentEventID = cruise) %>% 
  distinct(eventID, .keep_all = TRUE) %>% 
  mutate(eventRemarks = "station")

# Join date and coordinates to trawl. Remember that the date and coordinates are not from the Salmon Diet data sheet, but from the trawl data sheet. Minimum and maximum depth of the trawl be included here as well if the information becomes available. 
salmon_trawl <- salmon %>%
  left_join(select(
    trawl,
    station,
    UTC_start,
    X,
    Y
  )) %>%
  select(eventID = trawl,
         parentEventID = station,
         eventDate = UTC_start,
         decimalLatitude = Y,
         decimalLongitude = X) %>%
  distinct(eventID, .keep_all = TRUE) %>%
  mutate(eventRemarks = "trawl")

# The Salmon Diet data does not include information as to at what depths the trawls were conducted. Currently I believe this information is also not present in the general Trawl data. Should this information become available it should be connected to the general trawl information, along with info such as: volume water filter, speed at which the vessel was traveling during trawl, starting & finish lat/longs of trawl.
```

***

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 10px;}
</style>
<div class = "blue">
The current dilemma I have is that there is no unique `sampleID` associated with each data entry in the salmon diet data. This makes it difficult to properly structure the dataset. There are two options we can take:

Option A: If we want to treat a salmon as a `sampling instrument`, each salmon species should be in the Event Core, rather than in the Occurrence Core. This is a novel technique, I believe salmon haven't really been seen as an instrument to collect data. The measurement or facts (eMOF) associated to the salmon can be linked directly to the event layer. See the layout of option A [here](https://app.diagrams.net/#G1qMPIuqd90pT11CXULX6k48YjKS_5WDUx)

Option B: Alternatively, we create a unique sampleID for each data entry in the salmon diet data, but in that we do not regard the salmon as the sampling instrument (and, additionally, it would be difficult to associate any data to the sampleID). However, this way each salmon in the salmon diet data set would be an `occurrence`, with associated measurements or facts. See the layout of option B [here](https://app.diagrams.net/#G18hDRALexgAs5hQ8k89lsgg7SDI5Ho83H)
</div>

Currently the data is reformatted data is structured as shown in **option A**. However, the data standardization is an evolving process so the structure might change. 

***

``` {r, eval = FALSE}

# If we treat every salmon entry for which the stomach has been removed and examined as a sampling event, we essentially treat the salmon like a sampling instrument of their direct surroundings! 
 
salmon_sample <- salmon %>% 
  select(eventID = sample,
         parentEventID = trawl) %>% 
  distinct(eventID, .keep_all = TRUE) %>% 
  mutate(eventRemarks = "sample")

# To connect them all together: 
salmon_event <- bind_rows(salmon_cruise,salmon_station,salmon_trawl, salmon_sample) %>%
  select(eventID, parentEventID:decimalLongitude, eventRemarks) %>%
  mutate(type = "HumanObservation")

# Make sure the folder path exists already (e.g. ./Salmon Diet/tidy_data)
write_csv(salmon_event, here("Salmon Diet", "tidy_data", "salmon_event.csv"))
drive_upload(here("Salmon Diet", "tidy_data", "salmon_event.csv"),
             path = "https://drive.google.com/drive/u/0/folders/1pGd2cuvdb25QX7CGm2wC81Emh5O-LdbR",
             name = "salmon_event.csv",
             overwrite = TRUE)
```

# Occurrence Core

Currently we have opted for Option A, where the salmon caught and examined are in the Event Core, with a single Occurrence Core for the prey occurrences in the stomach content. This option was chosen as the focus of this particular study is on the stomach content, and to understand whether considering salmon as the `sampling instrument` is a viable option. 

The problem in the salmon diet data is that there are sometimes numerous individuals of the same species caught at a station/within a trawl, but these individuals have no defining tag. Additionally, some data entries cover multiple stomachs examined (`N stomachs` > 1), and the associated fork length and mass appear to be averaged across multiple individuals. However, in the current set up it's sometimes not possible to identify which individuals this covers (i.e., no mentioning of Floytags or #MGL).

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 10px;}
</style>
<div class = "blue">
How do the scientists know which individual the stomach content belong to, asides from their data?
</div>

The Occurrence core connected to each salmon stomach examined consists of the occurrences of prey species consumed by the salmon species at each station. The measurement attached to these unique occurrences is the mass (g) of each prey species consumed. Each prey species will receive their own unique URN from [WoRMS](http://www.marinespecies.org/), if identified.

As all the prey species are listed as columns, we first have to investigate all the unique prey species consumed, as these will get their associated WoRMS URN. There is also a column `digested_food` which will not have an associated URN, as it likely reflects unidentified prey species consumed -- this will have to be confirmed! This column will get added as a measurement linking directly to the salmon (Event Core). 

``` {r prey_occurrence, eval = FALSE}
salmon_prey <- salmon  %>%
  pivot_longer(cols = boreoteuthis_borealis:mictostomus_pacificus, 
               names_to = "scientificName", values_to = "Mass") 
salmon_prey <- salmon_prey[ !(salmon_prey$scientificName %in% "digested_food"),]
salmon_prey <- salmon_prey %>%
  mutate(occurrenceStatus = ifelse(is.na(salmon_prey$Mass), "absent", "present"),
         occurrenceID = paste(sample, "stomach_content", scientificName, sep = ":"),
         scientificNameID = case_when(
scientificName == "boreoteuthis_borealis" ~ "urn:lsid:marinespecies.org:taxname:342326",
# Boreoteuthis is not an accepted genus, Gonatopsis is the accepted version?
scientificName == "neocalanus_cristatus" ~ "urn:lsid:marinespecies.org:taxname:104470",
scientificName == "pisces" ~ "urn:lsid:marinespecies.org:taxname:1167",
scientificName == "tomopteris" ~ "urn:lsid:marinespecies.org:taxname:11676",
scientificName == "chuneola_major" ~ "urn:lsid:marinespecies.org:taxname:431286" ,
scientificName == "sergestes_pacificus" ~ " ",
# Is only the atlantic version accepted?
scientificName == "e_pacifica" ~ "urn:lsid:marinespecies.org:taxname:237851",
scientificName == "tarletonbeania_crenularis" ~ "urn:lsid:marinespecies.org:taxname:282927",
scientificName == "hyperia_sp" ~ "urn:lsid:marinespecies.org:taxname:101796",
scientificName == "clione_limacina" ~ "urn:lsid:marinespecies.org:taxname:139178",
scientificName == "th_spinifera" ~ "urn:lsid:marinespecies.org:taxname:237874",
# Assumption here is made that this concerns Thysanoessa spinifera
scientificName == "themisto_pacifica" ~ "urn:lsid:marinespecies.org:taxname:254453",
scientificName == "limacina_helicina" ~ "urn:lsid:marinespecies.org:taxname:140223",
scientificName == "primno_macropa" ~ "urn:lsid:marinespecies.org:taxname:180759",
scientificName == "th_inspinata" ~ "urn:lsid:marinespecies.org:taxname:237872",
# Assumption here is made that 'th' stands for Thysanoessa (genus)
scientificName == "coelenterata" ~ "urn:lsid:marinespecies.org:taxname:2",
# Coelenterata is a term encompassing the animal phyla Cnidaria and Ctenophora? 
# Therefore the lowest common rank is the kingdom Animalia
scientificName == "vanadis_sp" ~ "urn:lsid:marinespecies.org:taxname:129150",
scientificName == "sagitta_elegans" ~ "urn:lsid:marinespecies.org:taxname:105440",
# The accepted name is Parasagitta elegans
scientificName == "salpa_aspera" ~ "urn:lsid:marinespecies.org:taxname:137270",
scientificName == "cyclosalpa" ~ "urn:lsid:marinespecies.org:taxname:137227", 
scientificName == "th_longipes" ~ "urn:lsid:marinespecies.org:taxname:237873",
# Assumption is made that it is Thysanoessa longipes, Thysanoessa was the only genus in common between longipes, inspinata and spinifera. 
scientificName == "cephalopoda" ~ "urn:lsid:marinespecies.org:taxname:11707",
scientificName == "phronima_sedentaria" ~ "urn:lsid:marinespecies.org:taxname:103272",
scientificName == "oikopleura" ~ "urn:lsid:marinespecies.org:taxname:103367",
scientificName == "isopoda" ~ "urn:lsid:marinespecies.org:taxname:1131",
scientificName == "clio_pyramidata" ~ "urn:lsid:marinespecies.org:taxname:139033",
scientificName == "chormiphora_cucumis" ~ "urn:lsid:marinespecies.org:taxname:106382",
# Accepted name is Hormiphora cucumis?
scientificName == "symbolophorus_californiensis" ~ "urn:lsid:marinespecies.org:taxname:272733",
scientificName == "okutania_anonycha" ~ "urn:lsid:marinespecies.org:taxname:342285",
# Is the accepted name Berryteuthis anonychus?
scientificName == "sebastes" ~ "urn:lsid:marinespecies.org:taxname:126175",
scientificName == "mictostomus_pacificus" ~ "urn:lsid:marinespecies.org:taxname:274294")
# Spelling error? Accepted name is Microstomus pacificus?
) %>%
  rename(id = sample)

salmon_prey_occ <- salmon_prey %>%
  select(id, occurrenceID, scientificName, scientificNameID, occurrenceStatus) %>%
  mutate(basisOfRecord = "HumanObservation") %>%
  filter(occurrenceStatus == "present")

# Make sure the folder path exists already (e.g. ./Salmon Diet/tidy_data)
write_csv(salmon_prey_occ, here("Salmon Diet", "tidy_data", "salmon_prey_occ.csv"))
drive_upload(here("Salmon Diet", "tidy_data", "salmon_prey_occ.csv"),
             path = "https://drive.google.com/drive/u/0/folders/1pGd2cuvdb25QX7CGm2wC81Emh5O-LdbR",
             name = "salmon_prey_occ.csv",
             overwrite = TRUE)
```

If we choose option B then the following chunk of code is also needed to create the Occurrence Core specifically for the salmon caught and whose stomach contents have been examined: 

``` {r salmon_occ, eval = FALSE}
salmon_spp_occ <- salmon %>%
  complete(trawl, species, fill = list(n_stomachs = 0)) %>%
  rename(scientificName = species) %>% 
  mutate(occurrenceStatus = ifelse(n_stomachs > 0, "present", "absent"),
         occurrenceID = paste(trawl, scientificName, sep = ":"),
         scientificNameID = case_when(
scientificName == "Oncorhynchus gorbuscha" ~ "urn:lsid:marinespecies.org:taxname:127182",
scientificName == "Oncorhynchus kisutch" ~ "urn:lsid:marinespecies.org:taxname:127184",
scientificName == "Oncorhynchus keta" ~ "urn:lsid:marinespecies.org:taxname:127183",
scientificName == "Oncorhynchus nerka" ~ "urn:lsid:marinespecies.org:taxname:254569",
scientificName == "Oncorhynchus tschawytscha" ~ "urn:lsid:marinespecies.org:taxname:158075")) %>%
  # Accepted name has slightly different species name: O. tshawytscha
  rename(eventID = trawl) %>%
  select(eventID, occurrenceID, scientificName, scientificNameID, occurrenceStatus)

# Make sure the folder path exists already (e.g. ./Salmon Diet/tidy_data)
write_csv(salmon_spp_occ, here("Salmon Diet", "tidy_data", "salmon_spp_occ.csv"))
drive_upload(here("Salmon Diet", "tidy_data", "salmon_spp_occ.csv"),
             path = "https://drive.google.com/drive/u/0/folders/1pGd2cuvdb25QX7CGm2wC81Emh5O-LdbR",
             name = "salmon_spp_occ.csv",
             overwrite = TRUE)
```

# measurementOrFact Cores

So even though we (currently) have one Occurrence Core for Salmon Diet data, there are two measurement Or Fact tables, one linking directly to the salmon (Event) and the other linking to the stomach content (prey occurrence). Let's start with the first extended measurement Or Fact Core, where we associate measurements of `fork length`, `mass`, `food mass`, `gut fullness index (GFI)`, `N stomachs`, `CFI` and possibly `digested food (unidentified)` to the salmon examined.  

``` {r, eval = FALSE}
salmon_spp_measurement <- salmon %>%
  select(measurementID = sample,
         fl_cm:cfi_percent_adj,
         digested_food) %>%
  pivot_longer(cols = fl_cm:digested_food, 
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = case_when(
           measurementType == "fl_cm" ~ "FL",
           measurementType == "mass_g" ~ "mass",
           measurementType == "food_mass_g" ~ "food_mass",
           measurementType == "gfi_percent" ~ "GFI",
           measurementType == "n_stomachs" ~ "count",
           measurementType == "cfi_percent_adj" ~ "CFI",
           measurementType == "digested_food" ~ "digested_food"),
    measurementID = paste(measurementID, measurementType, sep = ":"),
    measurementTypeID = case_when(
    measurementType == "FL" ~ " ",
    # But standard length: 	http://vocab.nerc.ac.uk/collection/P01/current/SL01XX01/
    measurementType == "mass" ~ "http://vocab.nerc.ac.uk/collection/P24/current/NMASS/",
    measurementType == "food_mass" ~ "http://vocab.nerc.ac.uk/collection/P24/current/NMASS/",
    measurementType == "GFI" ~ " ",
    measurementType == "count" ~ " ",
    measurementType == "CFI" ~ " ",
    measurementType == "digested_food" ~ "http://vocab.nerc.ac.uk/collection/P24/current/NMASS/"),
         measurementUnit = case_when(
    measurementType == "FL" ~ "cm",
    measurementType == "mass" ~ "g",
    measurementType == "food_mass" ~ "g",
    measurementType == "GFI" ~ "percent",
    measurementType == "count" ~ "stomachs",
    # This seems off to me - want to find a better alternative.
    measurementType == "CFI" ~ "percent",
    measurementType == "digested_food" ~ "g"),
         measurementUnitID = case_when(
    measurementUnit == "cm" ~ "http://vocab.nerc.ac.uk/collection/P06/current/ULCM/",
    measurementUnit == "g" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UGRM/",
    measurementUnit == "percent" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UPCT/",
    measurementUnit == "stomachs" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UUUU/")) %>%
  select(measurementID, measurementType, measurementTypeID,
         measurementValue, measurementUnit, measurementUnitID)

# Write up csv file and upload to Google Drive folder
write_csv(salmon_spp_measurement, here("Salmon Diet", "tidy_data", "salmon_spp_measurement.csv"))
drive_upload(here("Salmon Diet", "tidy_data", "salmon_spp_measurement.csv"),
             path = "https://drive.google.com/drive/u/0/folders/1pGd2cuvdb25QX7CGm2wC81Emh5O-LdbR",
             name = "salmon_spp_eMoF.csv",
             overwrite = TRUE)
```

The second measurement Or Fact table will be created for the stomach content of the salmon species examined. The measurements associated to this Occurrence Core is `mass digested prey species`. First we filter out all the entries of species that were not recorded within a stomach (i.e., where measurementValue = NA)

``` {r, eval = FALSE}
salmon_prey_measurement <- salmon %>%
  select(measurementID = sample,
         boreoteuthis_borealis:mictostomus_pacificus) %>%
  select(-digested_food) %>%
  pivot_longer(cols = boreoteuthis_borealis:mictostomus_pacificus,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  filter(!is.na(measurementValue)) %>%
  mutate(measurementID = case_when(
    measurementType == "boreoteuthis_borealis" ~ paste(measurementID, "b.borealis", sep = ":"),
    measurementType == "neocalanus_cristatus" ~ paste(measurementID, "n.cristatus", sep = ":"),
    measurementType == "pisces" ~ paste(measurementID, "pisces", sep = ":"),
    measurementType == "tomopteris" ~ paste(measurementID, "tomopteris", sep = ":"),
    measurementType == "chuneola_major" ~ paste(measurementID, "c.major", sep = ":"),
    measurementType == "sergestes_pacificus" ~ paste(measurementID, "s.pacificus", sep = ":"),
    measurementType == "e_pacifica" ~ paste(measurementID, "e.pacifica", sep = ":"),
    measurementType == "tarletonbeania_crenularis" ~ paste(measurementID, "t.crenularis", sep = ":"),
    measurementType == "hyperia_sp" ~ paste(measurementID, "hyperia_sp", sep = ":"),
    measurementType == "clione_limacina" ~ paste(measurementID, "c.limacina", sep = ":"),
    measurementType == "th_spinifera" ~ paste(measurementID, "t.spinifera", sep = ":"),
    measurementType == "themisto_pacifica" ~ paste(measurementID, "t.pacifica", sep = ":"),
    measurementType == "limacina_helicina" ~ paste(measurementID, "l.helicina", sep = ":"),
    measurementType == "primno_macropa" ~ paste(measurementID, "p.macropa", sep = ":"),
    measurementType == "th_inspinata" ~ paste(measurementID, "t.inspinata", sep = ":"),
    measurementType == "coelenterata" ~ paste(measurementID, "coelenterata", sep = ":"),
    measurementType == "vanadis_sp" ~ paste(measurementID, "vanadis_sp", sep = ":"),
    measurementType == "sagitta_elegans" ~ paste(measurementID, "s.elegans", sep = ":"),
    measurementType == "salpa_aspera" ~ paste(measurementID, "s.aspera", sep = ":"),
    measurementType == "cyclosalpa" ~ paste(measurementID, "cyclosalpa", sep = ":"),
    measurementType == "th_longipes" ~ paste(measurementID, "t.longipes", sep = ":"),
    measurementType == "cephalopoda" ~ paste(measurementID, "cephalopoda", sep = ":"),
    measurementType == "phronima_sedentaria" ~ paste(measurementID, "p.sedentaria", sep = ":"),
    measurementType == "oikopleura" ~ paste(measurementID, "oikopleura", sep = ":"),
    measurementType == "isopoda" ~ paste(measurementID, "isopoda", sep = ":"),
    measurementType == "clio_pyramidata" ~ paste(measurementID, "c.pyramidata", sep = ":"),
    measurementType == "chormiphora_cucumis" ~ paste(measurementID, "c.cucumis", sep = ":"),
    measurementType == "symbolophorus_californiensis" ~ paste(measurementID, "s.californiensis", sep = ":"),
    measurementType == "okutania_anonycha" ~ paste(measurementID, "o.anonycha", sep = ":"),
    measurementType == "sebastes" ~ paste(measurementID, "sebastes", sep = ":"),
    measurementType == "mictostomus_pacificus" ~ paste(measurementID, "m.pacificus", sep = ":")),
  measurementUnit = "g",
  measurementUnitID = case_when(
    measurementUnit == "g" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UGRM/"
  )
  )

# Write up csv file and upload to Google Drive folder
write_csv(salmon_prey_measurement, here("Salmon Diet", "tidy_data", "salmon_prey_measurement.csv"))
drive_upload(here("Salmon Diet", "tidy_data", "salmon_prey_measurement.csv"),
             path = "https://drive.google.com/drive/u/0/folders/1pGd2cuvdb25QX7CGm2wC81Emh5O-LdbR",
             name = "salmon_prey_eMoF.csv",
             overwrite = TRUE)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 10px;}
</style>
<div class = "blue">

Should these measurementOrFact cores be combined into a single MeasuremtOrFact Core? 

</div>

