---
title: "data_wrangle_trawl2019"
author: "Tim van der Stap"
date: "4/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(tidyverse)
library(lubridate)
library(readxl)
library(here)
library(worrms)
library(obistools)
library(parsedate)
library(googledrive)
library(uuid)
```

The first step is to download the files from GoogleDrive and save it in a local folder. One file contains the bridgelog data and aggregated data, whereas the other contains specimen-specific catch measurements. 

``` {r file_download}
drive_download(file = "https://docs.google.com/spreadsheets/d/1Teh8Lcj3vCxocB2HPso19XYrZvCUspFt/edit#gid=1038314498", 
               path = here::here("Trawl", "2019", "raw_data", "2019_GoA_Fish_Trawl_catchdata.xlsx"),
               overwrite = TRUE)

drive_download(file = "https://docs.google.com/spreadsheets/d/1LxgxoGkZKa7Q_4XJ1XY7HOd4mhZ_epQz/edit#gid=2077327798",
               path = here::here("Trawl", "2019", "raw_data", "2019_GoA_Fish_Specimen_data.xlsx"),
               overwrite = TRUE)
```

The trawl data is read from the local folder, and the date/time is formatted to the ISO 8601 standards. Additionally, five layers are created to populate the Event Core: "project", "cruise", "station", "trawl" and "sample". Following the discussion [here](https://github.com/iobis/env-data/issues/10), data on total species' catch biomass gets linked to the `trawl` layer (through an `occurrenceID`) in the Event Core, whereas individual fish measurements get linked the `sample` layer (see diagram [here](https://lucid.app/lucidchart/02b19f65-6072-4633-b4a8-ea1557fdf51a/edit?page=0_0#)). The fish trawl abundance data file also includes the bridgelog file sheet, where metadata is recorded. When reading the data in from the local folder make sure that a clear distinction is made between the bridgelog (metadata) and the catch data.

``` {r trawl2019_data_prep, eval = FALSE, warning = FALSE}
trawl2019 <- read_excel(here("Trawl", "2019", "raw_data", 
                             "2019_GoA_Fish_Trawl_catchdata.xlsx"), sheet = "BRIDGE_LOG_FINAL")

# Time is recorded numerically (1037 instead of 10:37), so need to change these columns:
trawl2019$END_DEPLOYMENT_TIME <- substr(as.POSIXct(sprintf("%04.0f", trawl2019$END_DEPLOYMENT_TIME), format = "%H%M"), 12, 16)
trawl2019$BEGIN_RETRIEVAL_TIME <- substr(as.POSIXct(sprintf("%04.0f", trawl2019$BEGIN_RETRIEVAL_TIME), format = "%H%M"), 12, 16)

# Additionally, the vessel time is recorded in 'Vladivostok' according to the metadata tab. This has to be converted to UTC.  
trawl2019 <- trawl2019 %>%
  mutate(eventDate_start = format_iso_8601(as.POSIXct(paste(EVENT_DATE_START, END_DEPLOYMENT_TIME),
                                                      tz = "Asia/Vladivostok")),
         eventDate_start = str_replace(eventDate_start, "\\+00:00", "Z"),
         eventDate_finish = format_iso_8601(as.POSIXct(paste(EVENT_DATE_FINISH, BEGIN_RETRIEVAL_TIME),
                                                       tz = "Asia/Vladivostok")),
         eventDate_finish = str_replace(eventDate_finish, "\\+00:00", "Z"),
         eventDate = paste(eventDate_start, eventDate_finish, sep = "/"),
         project = "IYS",
         cruise = paste(project, "GoA2019", sep = ":"), 
         station = paste(cruise, TOW_NUMBER, sep="_Stn"),
         trawl = paste(station, "trawl", sep=":"))

# For OBIS it is useful to parse the eventDate and extract year, month and day from it, for visualizations purposes.
trawl2019$year <- as.numeric(format(as.Date(trawl2019$eventDate), "%Y"))
trawl2019$month <- as.numeric(format(as.Date(trawl2019$eventDate), "%m"))
trawl2019$day <- as.numeric(format(as.Date(trawl2019$eventDate), "%d"))

trawl2019_allCatch <- read_excel(here("Trawl", "2019", "raw_data", 
                                      "2019_GoA_Fish_Trawl_catchdata.xlsx"), sheet = "CATCH_FINAL") %>%
  mutate(project = "IYS",
         cruise = paste(project, "GoA2019", sep = ":"),
         station = paste(cruise, `TOW_NUMBER (number)`, sep = "_Stn"),
         trawl = paste(station, "trawl", sep = ":"))

trawl2019_specimen <- read_excel(here("Trawl", "2019", "raw_data", "2019_GoA_Fish_Specimen_data.xlsx"), 
                                 sheet = "SPECIMEN_FINAL") %>%
  mutate(project = "IYS",
         cruise = paste(project, "GoA2019", sep = ":"),
         station = paste(cruise, TOW_NUMBER, sep = "_Stn"),
         trawl = paste(station, "trawl", sep = ":"),
         sample = paste(trawl, "sample", sep = ":"),
         sample = paste(sample, row_number(), sep = ""))
```

***

## Event Core 

Next, dataframes are created for each unique `eventID` layer, with unique associated columns. Eventually, the data frames get joined, and the Event Core gets saved locally and in Google Drive. The column `type` is not a required column, but includes descriptive information pertaining to any issues that arose during the setting of the trawl net. 

``` {r trawl_event, eval = FALSE}
trawl2019_project <- trawl2019 %>%
  select(eventID = project) %>%
  distinct(eventID) %>%
  mutate(type = "project")

trawl2019_cruise <- trawl2019 %>% 
  select(eventID = cruise,
         parentEventID = project) %>% 
  distinct(eventID, .keep_all = TRUE) %>%
  mutate(type = "cruise")

trawl2019_station <- trawl2019 %>% 
  select(eventID = station,
         parentEventID = cruise) %>% 
  distinct(eventID, .keep_all = TRUE) %>%
  mutate(type = "station")

# The coordinates associated to the trawl need to be presented in a LINESTRING.
# END_LONGITUDE_DD needs to be inverted (has to be between -180 and 180, inclusive). 
trawl2019_coordinates <- trawl2019 %>%
  select(eventID = trawl,
         START_LATITUDE_DD,
         longitude,
         END_LATITUDE_DD,
         END_LONGITUDE_DD) %>%
  mutate(END_LONGITUDE_DD = END_LONGITUDE_DD * -1,
         footprintWKT = paste("LINESTRING (", longitude, START_LATITUDE_DD, ",", 
                              END_LONGITUDE_DD, END_LATITUDE_DD, ")")) 

trawl2019_linestring <- obistools::calculate_centroid(trawl2019_coordinates$footprintWKT)
trawl2019_linestring <- cbind(trawl2019_coordinates, trawl2019_linestring) %>%
  select(eventID, footprintWKT, decimalLatitude, decimalLongitude, coordinateUncertaintyInMeters)

trawl2019_trawl <- trawl2019 %>% 
  select(eventID = trawl,
         parentEventID = station,
         eventDate,
         year,
         month,
         day,
         fieldNotes = PROBLEM_DESCRIPTION) %>%
  mutate(minimumDepthInMeters = 0, # trawl was at the surface
         maximumDepthInMeters = max(trawl2019$MOUTH_OPENING_WIDTH),
         samplingProtocol = "surface trawl", # when available add DOI to paper here
         locality = case_when(
           trawl2019$EVENT_SUB_TYPE == "Can EEZ" ~ "Canadian EEZ"),
         locationID = case_when(
           trawl2019$EVENT_SUB_TYPE == "Can EEZ" ~ "http://marineregions.org/mrgid/8493")) %>%
  left_join(trawl2019_linestring, by = "eventID") %>% 
  distinct(eventID, .keep_all = TRUE) %>%
    mutate(type = "surface trawl")
  
trawl2019_sample <- trawl2019_specimen %>%
  select(eventID = sample,
         parentEventID = trawl) %>%
  distinct(eventID, .keep_all = TRUE) %>%
  mutate(type = "individual sample")

trawl2019_event <- bind_rows(trawl2019_project, 
                             trawl2019_cruise,
                             trawl2019_station,
                             trawl2019_trawl,
                             trawl2019_sample) 

col_order <- c("eventID", "parentEventID", "eventDate", "year", "month", "day", "decimalLatitude", "decimalLongitude", "footprintWKT", "coordinateUncertaintyInMeters", "minimumDepthInMeters", "maximumDepthInMeters", "samplingProtocol", "locality", "locationID", "type", "fieldNotes")
trawl2019_event <- trawl2019_event[, col_order]

# Flatten event records:
trawl2019_event <- obistools::flatten_event(trawl2019_event)

# Save locally and on Google Drive:
write_csv(trawl2019_event, here("Trawl", "2019", "tidy_data", "trawl2019_event.csv"))
drive_upload(here("Trawl", "2019", "tidy_data", "trawl2019_event.csv"),
             path = "",
             name = "trawl2019_eventCore.csv",
             overwrite = TRUE)
```

## Occurrence

As the trawl data consists of biotic data, an occurrence extension table is created, whereby unique `occurrenceIDs` get linked to the `eventID`. The `occurrenceID` is created from, and nested under, the `eventID`, and any measurements pertaining to the overall catch _or_ specific individual will be associated to that `occurrenceID` in the extended MeasurementOrFact (eMOF) extension.

The Occurrence extension will consist of two tables merged, one from the aggregated (summary) catch data (`catch_final`), and one from the individual length and weight measurements. There will be unique `occurrenceIDs` associated to the trawl layer in Event Core. Measurements associated with this `occurrenceID` will be total species' catch biomass and total number caught. The `occurrenceID` for the aggregated catch data will have the format `IYS:GoA2019_Stn1:trawl:occ-1`. An additional layer in the Event Core (sample) was created so that other unique `occurrenceIDs` can be linked to the individual level, with `occurrenceIDs` having the format `IYS:GoA2019_Stn1:trawl:sample1:occ-1`. Measurements linked to these `occurrenceIDs` are individual length and weight. Once more data becomes available for individual fish specimen, these will be added at a later stage. For a diagram on the relational database structure, see [here](https://lucid.app/lucidchart/02b19f65-6072-4633-b4a8-ea1557fdf51a/edit?page=0_0#).

Additionally, a `resourceRelationship` extension will be created that will allow data users to understand that some `occurrenceID`s will be a subset ('located within') of the aggregated data. 

For discussion on how `occurrenceIDs` are linked to different layers, see [here](https://github.com/iobis/env-data/issues/10).

A required term (column) for the Occurrence extension is `occurrenceStatus`, which is used to define presence/absence of a taxonomic record. The package `worrms` is used to generate hierarchical taxonomic information on the species observed. For the `scientificNameID`, the package pulls the Life Science Identifier (LSID) from the WoRMS database (http://www.marinespecies.org/). 

Finally, we can do a quick check to make sure that all the unique species are included in the Occurrence extension dataframe. Once confirmed that all species listed and recorded in the original Processed Data are also included in the Standardized Data, the Occurrence extension can be saved locally and in the correct Google Drive folder. We also need to make sure that _if_ columns 'SPECIES' or 'SPECIES (Description if needed)' make mention of a lifestage, or sex, we need to include this in the occurrence extension under separate columns. Please be aware that the column `species` in DwC terms is referred to as the `specificEpithet`.

``` {r trawl2019_occ, eval = FALSE}
trawl2019_allCatch <- trawl2019_allCatch %>% 
  mutate(scientificname = paste(genus, species, sep = " "))

# The scientificname includes a lot of 'sp.', which often is not read in properly by the worrms package. Therefore, this has to be removed, and included as an identificationQualifier. Once included in the identificationQualifier, remove from scientificname. 
trawl2019_allCatch$identificationQualifier <- ifelse(grepl("\\b sp.\\b|\\b sp\\b", trawl2019_allCatch$scientificname),
                                                     "sp. inc.", NA)
trawl2019_allCatch$scientificname <- gsub("\\b sp.\\b|\\b sp\\b", "", trawl2019_allCatch$scientificname)
trawl2019_allCatch$species <- gsub("\\bsp.\\b|\\bsp\\b", NA, trawl2019_allCatch$species)
         
trawl2019_allCatch_sn <- worrms::wm_records_names(unique(trawl2019_allCatch$scientificname)) %>% bind_rows()
trawl2019_allCatch_taxa <- left_join(trawl2019_allCatch, trawl2019_allCatch_sn, by = "scientificname")

# You'll notice that not all unique_taxa have associated taxonomic hierarchial information. Let's inspect these observations as they might be either misspelled, or a non-Latin name is provided. 
unidentified_species <- trawl2019_allCatch_taxa %>% filter(is.na(AphiaID)) %>% distinct(scientificname)
```

A total of 6 species could not be matched to a record within the WoRMS database. Most often this is due to spelling errors, the inclusion of '.sp', or the lack of supplying a Latin name. I manually check these species in the WoRMS database and confirm with the data provider that the suggested species are the ones meant. These species will then be corrected in the trawl2019_allCatch_taxa data frame. 

- [ ] TO DO: Confirm corrections below with Chrys Neville. 

``` {r trawl2019_occ cont, eval = FALSE}
trawl2019_allCatch$scientificname <- gsub("Phacellophora camtshchatica", "Phacellophora camtschatica", 
                                          trawl2019_allCatch$scientificname)
trawl2019_allCatch$scientificname <- gsub("Ctenephores", "Ctenophora", trawl2019_allCatch$scientificname)
trawl2019_allCatch$scientificname <- gsub("Moroteuthis robusta", "Moroteuthis robustus", trawl2019_allCatch$scientificname)
trawl2019_allCatch$scientificname <- gsub("Chrysaora melonaster", "Chrysaora melanaster", trawl2019_allCatch$scientificname)
trawl2019_allCatch$scientificname <- gsub("Paralepididae gen. sp. 1", "Paralepididae", trawl2019_allCatch$scientificname)
trawl2019_allCatch$scientificname <- gsub("NA Siphonophora", "Siphonophora", trawl2019_allCatch$scientificname)
  
# So now we can run worrms::wm_records_names again, and there should be no NA in the AphiaID column.
trawl2019_allCatch_worms <- worrms::wm_records_names(unique(trawl2019_allCatch$scientificname)) %>% bind_rows()
trawl2019_occ <- left_join(trawl2019_allCatch, trawl2019_allCatch_worms, by = "scientificname") %>%
  rename(eventID = trawl,
         specificEpithet = species,
         scientificNameAuthorship = authority,
         taxonomicStatus = status,
         taxonRank = rank,
         scientificName = scientificname,
         scientificNameID = lsid,
         individualCount = `CATCH_COUNT (pieces)(**includes Russian expansion for some species)`) %>%
  mutate(occurrenceID = paste(eventID, "occ", sep = ":"),
         occurrenceID = paste(occurrenceID, row_number(), sep = "-"),
         occurrenceStatus = "present",
         lifeStage = NA, 
         sex = NA)

# By joining those two dataframes, the column 'genus' is duplicated. Make sure to select the proper genus column, the one generated from the worrms package: 
trawl2019_occ_fnl <- trawl2019_occ %>%
  select(eventID, occurrenceID, occurrenceStatus,
         specificEpithet, scientificName, scientificNameID, individualCount,
         identificationQualifier, scientificNameAuthorship, taxonomicStatus, 
         taxonRank, kingdom, phylum, class, order, family, genus.y, lifeStage, sex)
```

So now we have created the first Occurrence extension table, which comes from the overall catch data. Please note that this extension does _not_ include information on count, this will be added in the extended MeasurementOrFact (eMOF) extension. In the next section, we create a second occurrence extension table, with `occurrenceID`s created for each unique individual whose length has been measured. This information will come from the `2019_GoA_Fish_Speciment_data.xlsx` file. 

Once the Occurrence extensions are combined, a column will have to be added to indicate that the individual taxonomic records (as presented in the `trawl2019_catch_ind` dataframe) are a subset of the overall catch at that station. We will use the column `occurrenceRemarks` for this.  

``` {r occ_species, eval = FALSE}
trawl2019_catch_ind <- trawl2019_specimen %>%
  rename(eventID = sample) %>%
  mutate(scientificname = paste(genus, species, sep = " "))

trawl2019_catch_ind_sn <- worrms::wm_records_names(unique(trawl2019_allCatch$scientificname)) %>% bind_rows()
trawl2019_catch_ind_taxa <- left_join(trawl2019_catch_ind, trawl2019_catch_ind_sn, by = "scientificname")

# Lets inspect whether there are any unidentified species:
unidentified_species_ind <- trawl2019_catch_ind_taxa %>% filter(is.na(AphiaID)) %>% distinct(scientificname)

trawl2019_catch_ind$identificationQualifier <- ifelse(grepl("\\b sp.\\b|\\b sp. 1\\b", trawl2019_catch_ind$scientificname),
                                                     "sp. inc.", NA)
trawl2019_catch_ind$scientificname <- gsub("\\b sp.\\b|\\b sp. 1\\b", "", trawl2019_catch_ind$scientificname)

# Make minor adjustments to recorded species name to ensure match with WoRMS database:
trawl2019_catch_ind$scientificname <- gsub("Moroteuthis robusta", "Moroteuthis robustus", trawl2019_catch_ind$scientificname)
trawl2019_catch_ind$scientificname <- gsub("Chrysaora melonaster", "Chrysaora melanaster", trawl2019_catch_ind$scientificname)

trawl2019_catch_ind$specificEpithet <- stringr::word(trawl2019_catch_ind$scientificname, 2)

# So now we can run worrms::wm_records_names again, and there should be no NA in the AphiaID column.
trawl2019_catch_ind_worms <- worrms::wm_records_names(unique(trawl2019_catch_ind$scientificname)) %>% bind_rows()

trawl2019_catch_ind_occ <- left_join(trawl2019_catch_ind, trawl2019_catch_ind_worms, by = "scientificname") %>%
  rename(scientificNameAuthorship = authority,
         taxonomicStatus = status,
         taxonRank = rank,
         scientificName = scientificname,
         scientificNameID = lsid) %>%
  mutate(occurrenceID = paste(eventID, "occ", sep = "-"),
         occurrenceStatus = "present",
         individualCount = 1)

# By joining those two dataframes, the column 'genus' is duplicated. Make sure to select the proper genus column, the one generated from the worrms package: 
trawl2019_catch_ind_fnl <- trawl2019_catch_ind_occ %>%
  select(eventID, occurrenceID, occurrenceStatus, scientificName, scientificNameID, individualCount,
         specificEpithet, scientificNameAuthorship, taxonomicStatus, taxonRank, kingdom, phylum, 
         class, order, family, genus.y, identificationQualifier, SEX_CODE)

# sex and lifestage is recorded in the SEX_CODE column - information in this column needs to be split out in separate columns matching DwC terminology: 
trawl2019_catch_ind_fnl <- trawl2019_catch_ind_fnl %>%
  mutate(sex = case_when(
    SEX_CODE == "M" ~ "male",
    SEX_CODE == "F" ~ "female",
    SEX_CODE == "N" ~ "unknown"),
         lifeStage = case_when(
    SEX_CODE == "J" ~ "juvenile",
    SEX_CODE == "3" ~ " ")) %>%
  select(-SEX_CODE)
```

- [ ] TO DO: Ask Chrys Neville what SEX_CODE = 3 implies. 

To finalize the Occurrence extension, combine the two occurrence tables. I prefer to re-order it so that it visually makes sense to me. Then, save the combined Occurrence extension locally and in Google Drive. 

``` {r trawl occurrence final, eval = FALSE}
trawl2019_occ_ext <- dplyr::bind_rows(trawl2019_occ_fnl, trawl2019_catch_ind_fnl)

# To re-order the occurrenceID, use following code:
order <- stringr::str_sort(trawl2019_occ_ext$occurrenceID, numeric=TRUE)
trawl2019_occ_ext <- trawl2019_occ_ext[match(order, trawl2019_occ_ext$occurrenceID),]

# Save the occurrence Core locally and in Google Drive:
write_csv(trawl2019_occ_ext, here("Trawl","2019", "tidy_data", "trawl2019_occ_final.csv"))
```

***

## extended MeasurementOrFact (eMOF) Extension

We create a few measurements or fact tables initially - one for the measurements pertaining to the overall catch (catch_weight and total catch numbers), and one with the measurements pertaining to selected individuals (individual length and weight). These tables will eventually be merged into a single extended measurementOrFact (eMOF) extension. 

Additionally, we need to create eMOF tables with measurements pertaining to the sampling event: tow speed, etc. These come from the `trawl2019` dataset. Variables in the samplingEffort category are `tow_speed`, `tow_duration`, `tow_distance` and `tow_direction`. There is also data collected on bottom depth, every 15 minutes. We save these tables separately, and also combine them to form one final eMOF extension. Through the IPT, multiple eMOF extensions can be uploaded if preferred. Including them all in a single extension might make it too big. 

Following the [trawl diagram](https://lucid.app/lucidchart/02b19f65-6072-4633-b4a8-ea1557fdf51a/edit?page=0_0#), information on the research vessel gets added to the `cruise` layer in the Event Core:

``` {r research_vessel, eval = FALSE}
trawl2019_rv <- trawl2019 %>%
  select(eventID = cruise) %>%
  mutate_all(as.character) %>%
  mutate(measurementType = "Research Vessel",
         measurementTypeID = "http://vocab.nerc.ac.uk/collection/L06/current/31/",
         measurementValue = "R/V Prof. Kaganovskiy",
         measurementValueID = "http://vocab.nerc.ac.uk/collection/C17/current/90ZH/",
         measurementUnit = NA,
         measurementUnitID = NA) %>% distinct()
```

Next, information on the bottom depth throughout the trawl is added: 

```{r bottomdepth, eval = FALSE}
trawl2019_bottomdepth <- trawl2019 %>%
  select(eventID = trawl,
         BOTTOM_DEPTH_START_METERS,
         BOTTOM_DEPTH_END_METERS) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = BOTTOM_DEPTH_START_METERS:BOTTOM_DEPTH_END_METERS,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
                                  BOTTOM_DEPTH_START_METERS = "bottom_depth_trawl_start",
                                  BOTTOM_DEPTH_END_METERS = "bottom_depth_trawl_finish"),
         measurementTypeID = case_when(
           measurementType == "bottom_depth_trawl_start" ~ "http://vocab.nerc.ac.uk/collection/A05/current/EV_BATHY/",
           measurementType == "bottom_depth_trawl_finish" ~ "http://vocab.nerc.ac.uk/collection/A05/current/EV_BATHY/"),
         measurementUnit = "meters"),
         measurementUnitID = case_when(
           measurementUnit == "meters" ~ "http://vocab.nerc.ac.uk/collection/P06/current/ULAA/"),
         measurementValueID = NA,
         measurementID = paste(eventID, measurementType, sep = ":")) %>%
  select(eventID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)
# Save the bottom depth extension locally:
write_csv(trawl_bottomdepth, here("Trawl", "tidy_data", "trawl2020_bottomdepth_emof.csv"))
```

Measurements pertaining to the sampling effort: 

```{r samplingEffort, eval = FALSE}
trawl2019_samplingEffort <- trawl2019 %>%
  select(eventID = trawl,
         TOW_DURATION,
         SPEED,
         DISTANCE_TRAVELLED,
         BEARING) %>%
  mutate(TOW_DURATION = round(TOW_DURATION * 60)) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = TOW_DURATION:BEARING,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
         TOW_DURATION = "tow_duration",
         BEARING = "direction_of_set",
         SPEED = "vessel_speed",
         DISTANCE_TRAVELLED = "tow_distance"),
         measurementTypeID = case_when(
           measurementType == "tow_duration" ~ "http://vocab.nerc.ac.uk/collection/P01/current/AZDRZZ01/",
           measurementType == "direction_of_set" ~ "http://vocab.nerc.ac.uk/collection/P07/current/CFSN0481/",
           measurementType == "vessel_speed" ~ "http://vocab.nerc.ac.uk/collection/P01/current/TOWSPEED/",
           measurementType == "tow_distance" ~ "http://vocab.nerc.ac.uk/collection/P01/current/LENTRACK/"),
         measurementUnit = case_when(
           measurementType == "tow_duration" ~ "minutes",
           measurementType == "direction_of_set" ~ "degrees",
           measurementType == "vessel_speed" ~ "knots",
           measurementType == "tow_distance" ~ "nautical miles"),
         measurementUnitID = case_when(
           measurementUnit == "minutes" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UMIN/",
           measurementUnit == "degrees" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UABB/",
           measurementUnit == "knots" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UKNT/",
           measurementUnit == "nautical miles" ~ " "),
         measurementValueID = NA,
         measurementID = paste(eventID, measurementType, sep = ":")) %>%
  select(eventID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)

# Save the extended measurementOrFact extension locally and in Google Drive:
write_csv(trawl2019_samplingEffort, here("Trawl", "tidy_data", "trawl2020_samplingEffort_emof.csv"))
```

And chunk of code to assign sampling instrument measurements/facts:

```{r samplingInstrument, eval = FALSE}
trawl2019_samplingInstrument <- trawl2019 %>%
  select(eventID = trawl,
         GEAR_CODE,
         MOUTH_OPENING_WIDTH,
         MOUTH_OPENING_HEIGHT) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = GEAR_CODE:MOUTH_OPENING_HEIGHT,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
         GEAR_CODE = "sampling_gear",
         MOUTH_OPENING_WIDTH = "net_opening_width",
         MOUTH_OPENING_HEIGHT = "net_opening_height"),
         measurementTypeID = case_when(
           measurementType == "sampling_gear" ~ "http://vocab.nerc.ac.uk/collection/Q01/current/Q0100002/",
           measurementType == "net_opening_width" ~ "http://vocab.nerc.ac.uk/collection/Q01/current/Q0100014/",
           measurementType == "net_opening_height" ~ "http://vocab.nerc.ac.uk/collection/Q01/current/Q0100013/"),
         measurementUnit = case_when(
           measurementType != "sampling_gear" ~ "meters"),
         measurementUnitID = case_when(
           measurementUnit == "meters" ~ "http://vocab.nerc.ac.uk/collection/P06/current/ULAA/"),
         measurementValueID = case_when(
           measurementType == "sampling_gear" ~ " "),
         measurementID = paste(eventID, measurementType, sep = ":")) %>%
  select(eventID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)

# Save the extended measurementOrFact extension locally and in Google Drive:
write_csv(trawl2019_samplingInstrument, here("Trawl", "tidy_data", "trawl2020_samplingInstrument_emof.csv"))
```

And a chunk of code to assign environmental conditions during the trawl:

```{r environmental conditions, eval = FALSE}
trawl2019_conditions <- trawl2019 %>%
  select(eventID = trawl,
         WIND_DIRECTION,
         WIND_SPEED,
         SEA_STATE_ID,
         WATER_TEMPERATURE) %>%
  mutate(SWELL_HEIGHT = ifelse(grepl("swell", trawl2019_conditions$SEA_STATE_ID), trawl2019_conditions$SEA_STATE_ID, NA),
         SWELL_HEIGHT = gsub("swell ", "", trawl2019_conditions$SWELL_HEIGHT),
         SEA_STATE_ID = ifelse(grepl("swell", trawl2019_conditions$SEA_STATE_ID), NA, trawl2019_conditions$SEA_STATE_ID)) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = WIND_DIRECTION:SWELL_HEIGHT,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
                                  SEA_STATE_ID = "wave_height",
                                  WATER_TEMPERATURE = "water_temperature",
                                  WIND_DIRECTION = "wind_direction",
                                  WIND_SPEED = "wind_speed",
                                  SWELL_HEIGHT = "swell_height"),
         measurementTypeID = case_when(
           measurementType == "wave_height" ~ "http://vocab.nerc.ac.uk/collection/P07/current/JNQS0CMX/",
           measurementType == "swell_height" ~ "http://vocab.nerc.ac.uk/collection/P09/current/SWHT/",
           measurementType == "water_temperature" ~ "http://vocab.nerc.ac.uk/collection/P01/current/TEMPPR01/",
           measurementType == "wind_direction" ~ "http://vocab.nerc.ac.uk/collection/B39/current/relwinddir/",
           measurementType == "wind_speed" ~ "http://vocab.nerc.ac.uk/collection/P07/current/CFSN0038/"),
         measurementUnit = case_when(
           measurementType == "wave_height" ~ "meters",
           measurementType == "swell_height" ~ "meters",
           measurementType == "wind_direction" ~ "degrees",
           measurementType == "wind_speed" ~ "knots",
           measurementType == "water_temperature" ~ "degrees celsius"),
         measurementUnitID = case_when(
           measurementUnit == "meters" ~ "http://vocab.nerc.ac.uk/collection/P06/current/ULAA/",
           measurementUnit == "degrees" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UAAA/",
           measurementUnit == "knots" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UKNT/",
           measurementUnit == "degrees celsius" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UPAA/"),
         measurementValueID = NA,
         measurementID = paste(eventID, measurementType, sep = ":")) %>% 
  drop_na(measurementValue) %>%
  select(eventID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)
  

# Save the extended measurementOrFact extension locally and in Google Drive:
write_csv(trawl2019_conditions, here("Trawl", "tidy_data", "trawl2019_conditions_emof.csv"))
```

Next two eMOF extensions are created with measurements pertaining to overall species' catch weight at each trawl, and individual length and weight measurements.

```{r eMoF}
trawl2019_allcatch_measurement <- trawl2019_occ %>%
  select(occurrenceID,
         individualCount,
         Mass_net,
         `length min (lmin)`,
         `length max (lmax)`,
         `age group ( for salmon 5= juv, 4 immature, 3 mature, 0 unknown)`) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = individualCount:`age group ( for salmon 5= juv, 4 immature, 3 mature, 0 unknown)`,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
                                  Mass_net = "Total species biomass in net",
                                  `length min (lmin)` = "length_min",
                                  `length max (lmax)` = "length_max",
                                  `age group ( for salmon 5= juv, 4 immature, 3 mature, 0 unknown)` = "age_group"),
         measurementID = case_when(
           measurementType == "individualCount" ~ paste(occurrenceID, "individualCount", sep = "-"),
           measurementType == "Total species biomass in net" ~ paste(occurrenceID, "biomass", sep = "-"),
           measurementType == "length_min" ~ paste(occurrenceID, "length_min", sep = "-"),
           measurementType == "length_max" ~ paste(occurrenceID, "length_max", sep = "-"),
           measurementType == "age_group" ~ paste(occurrenceID, "age_group", sep = "-")),
         measurementTypeID = case_when(
           measurementType == "Total species biomass in net" ~ "http://vocab.nerc.ac.uk/collection/S06/current/S0600088/",
           measurementType == "individualCount" ~ "http://vocab.nerc.ac.uk/collection/P01/current/OCOUNT01/",
           measurementType == "length_min" ~ "http://vocab.nerc.ac.uk/collection/P01/current/OBSMINLX/",
           measurementType == "length_max" ~ "http://vocab.nerc.ac.uk/collection/P01/current/OBSMAXLX/",
           measurementType == "age_group" ~ "http://vocab.nerc.ac.uk/collection/P01/current/LSTAGE01/"),
         measurementUnit = case_when(
           measurementType == "Total species biomass in net" ~ "kilogram",
           measurementType == "individualCount" ~ "individuals",
           measurementType == "length_min" ~ "centimeters",
           measurementType == "length_max" ~ "centimeters"),
         measurementUnitID = case_when(
           measurementUnit == "kilogram" ~ "http://vocab.nerc.ac.uk/collection/P06/current/KGXX/",
           measurementUnit == "individuals" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UUUU/",
           measurementUnit == "centimeters" ~ "http://vocab.nerc.ac.uk/collection/P06/current/ULCM/"),
         measurementValueID = case_when(
           measurementValue == "0" ~ "http://vocab.nerc.ac.uk/collection/S11/current/S1131/",
           measurementValue == "1" ~ " ",
           measurementValue == "3" ~ "http://vocab.nerc.ac.uk/collection/S11/current/S1116/",
           measurementValue == "4" ~ "http://vocab.nerc.ac.uk/collection/S11/current/S1171/",
           measurementValue == "5" ~ "http://vocab.nerc.ac.uk/collection/S11/current/S1127/",
           measurementValue == "3,4" ~ "http://vocab.nerc.ac.uk/collection/S11/current/S1145/",
           measurementValue == "3,4,5" ~ "http://vocab.nerc.ac.uk/collection/S11/current/S1145/",
           measurementValue == "4,5" ~ "http://vocab.nerc.ac.uk/collection/S11/current/S1127/"
         )) %>%
  select(occurrenceID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)
```

Next, create the measurementOrFact extension with length and weight measurements assigned to the _individual_ specimens. Some changes here are made in the Processed Data: The `SEX_CODE` covers both sex (male, female, unknown) and lifestage (J - which is juvenile? or should it be i for immature?)

- [ ] TO DO: Ask Chrys Neville which species were measured what way (total, fork, mantle length). 

``` {r}
trawl2019_ind_emof <- trawl2019_catch_ind_occ %>%
  select(occurrenceID,
         LENGTH, `STANDARD LENGTH`, SEX_CODE, `WEIGHT - TOTAL`) %>%
  mutate(lifeStage = ifelse(grepl("3|J", trawl2019_ind_emof$SEX_CODE), trawl2019_ind_emof$SEX_CODE, NA),
         SEX_CODE = gsub("3|J", NA, trawl2019_ind_emof$SEX_CODE)) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = LENGTH:`WEIGHT - TOTAL`,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
                                  LENGTH = "species length",
                                  `STANDARD LENGTH` = "Standard length",
                                  SEX_CODE = "sex",
                                  `WEIGHT - TOTAL` = "total weight"),
         measurementID = case_when(
           measurementType == "species length" ~ paste(occurrenceID, "species_length", sep = "-"),
           measurementType == "Standard length" ~ paste(occurrenceID, "standard_length", sep = "-"),
           measurementType == "sex" ~ paste(occurrenceID, "sex", sep = "-"),
           measurementType == "lifeStage" ~ paste(occurrenceID, "lifeStage", sep = "-"),
           measurementType == "total weight" ~ paste(occurrenceID, "total_weight", sep = "-")),
         measurementTypeID = case_when(
           measurementType == "species length" ~ "http://vocab.nerc.ac.uk/collection/P01/current/OBSINDLX/",
           measurementType == "Standard length" ~ "http://vocab.nerc.ac.uk/collection/P01/current/SL01XX01/",
           measurementType == "sex" ~ "http://vocab.nerc.ac.uk/collection/P01/current/ENTSEX01/",
           measurementType == "lifeStage" ~ "http://vocab.nerc.ac.uk/collection/P01/current/LSTAGE01/",
           measurementType == "total weight" ~ "http://vocab.nerc.ac.uk/collection/S06/current/S0600088/"),
         measurementUnit = case_when(
           measurementType == "species length" ~ "millimeters",
           measurementType == "Standard length" ~ "millimeters",
           measurementType == "total weight" ~ "grams"),
         measurementUnitID = case_when(
           measurementUnit == "millimeters" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UXMM/",
           measurementUnit == "grams" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UGRM/"),
         measurementValueID = case_when(
           measurementValue == "F" ~ "http://vocab.nerc.ac.uk/collection/S10/current/S102/"
           measurementValue == "M" ~ "http://vocab.nerc.ac.uk/collection/S10/current/S103/",
           measurementValue == "N" ~ "http://vocab.nerc.ac.uk/collection/S10/current/S104/",
           measurementValue == "J" ~ "http://vocab.nerc.ac.uk/collection/S11/current/S1127/",
           measurementValue == "3" ~ "http://vocab.nerc.ac.uk/collection/S11/current/S1116/")) %>%
  # Assuming the "3" stands for mature, similar to description for catch_final data. 
  mutate(measurementValueID = NA) %>%
  drop_na(measurementValue) %>%
  select(occurrenceID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)
```

Combine the two extended measurementOrFact extensions into a single eMOF extension, and save locally and in GoogleDrive:

``` {r}
trawl2020_eMOF <- dplyr::bind_rows(trawl2020_allcatch_emof, trawl2020_ind_emof)
# To combine all the extensions, use: 
# trawl2020_all_emofs <- dplyr::bind_rows(trawl_bottomdepth, samplingEffort, samplingInstrument, trawl_conditions, trawl2020_allcatch_emof, trawl2020_ind_emof)
write_csv(trawl2020_eMOF, here("Trawl", "tidy_data", "trawl_eMOF.csv"))
drive_upload(here("Trawl", "tidy_data", "trawl_emof.csv"),
             path = "https://drive.google.com/drive/folders/1ZbvwjD8odJr2SUKq6Z8zAwpLd6QeOOYC",
             name = "trawl_eMoF.csv",
             overwrite = TRUE)
```