---
title: "data_wrangle_trawl2019"
author: "Tim van der Stap"
date: "4/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(tidyverse)
library(lubridate)
library(readxl)
library(here)
library(worrms)
library(obistools)
library(parsedate)
library(googledrive)
library(uuid)
```

The first step is to download the files from GoogleDrive and save it in a local folder. One file contains the bridgelog data and aggregated data, whereas the other contains specimen-specific catch measurements. 

``` {r file_download}
drive_download(file = "https://docs.google.com/spreadsheets/d/1Teh8Lcj3vCxocB2HPso19XYrZvCUspFt/edit#gid=1038314498", 
               path = here::here("Trawl", "2019", "raw_data", "2019_GoA_Fish_Trawl_catchdata.xlsx"),
               overwrite = TRUE)

drive_download(file = "https://docs.google.com/spreadsheets/d/1LxgxoGkZKa7Q_4XJ1XY7HOd4mhZ_epQz/edit#gid=2077327798",
               path = here::here("Trawl", "2019", "raw_data", "2019_GoA_Fish_Specimen_data.xlsx"),
               overwrite = TRUE)
```

The trawl data is read from the local folder, and the date/time is formatted to the ISO 8601 standards. Additionally, five layers are created to populate the Event Core: "project", "cruise", "station", "trawl" and "sample". Following the discussion [here](https://github.com/iobis/env-data/issues/10), data on total species' catch biomass gets linked to the `trawl` layer (through an `occurrenceID`) in the Event Core, whereas individual fish measurements get linked the `sample` layer (see diagram [here](https://lucid.app/lucidchart/02b19f65-6072-4633-b4a8-ea1557fdf51a/edit?page=0_0#)). The fish trawl abundance data file also includes the bridgelog file sheet, where metadata is recorded. When reading the data in from the local folder make sure that a clear distinction is made between the bridgelog (metadata) and the catch data.

``` {r trawl2019_data_prep, eval = FALSE, warning = FALSE}
trawl2019 <- read_excel(here("Trawl", "2019", "raw_data", 
                             "2019_GoA_Fish_Trawl_catchdata.xlsx"), sheet = "BRIDGE_LOG_FINAL")

# Time is recorded numerically (1037 instead of 10:37), so need to change these columns:
trawl2019$END_DEPLOYMENT_TIME <- substr(as.POSIXct(sprintf("%04.0f", trawl2019$END_DEPLOYMENT_TIME), format = "%H%M"), 12, 16)
trawl2019$BEGIN_RETRIEVAL_TIME <- substr(as.POSIXct(sprintf("%04.0f", trawl2019$BEGIN_RETRIEVAL_TIME), format = "%H%M"), 12, 16)

# Not sure if the following code is the most efficient:
trawl2019 <- trawl2019 %>%
  mutate(eventDate_start = paste(EVENT_DATE_START, END_DEPLOYMENT_TIME),
         eventDate_start = str_replace(eventDate_start, " ", "T"),
         eventDate_start = paste0(eventDate_start, "Z"),
         eventDate_finish = paste(EVENT_DATE_FINISH, BEGIN_RETRIEVAL_TIME),
         eventDate_finish = str_replace(eventDate_finish, " ", "T"),
         eventDate_finish = paste0(eventDate_finish, "Z"),
         eventDate = paste(eventDate_start, eventDate_finish, sep = "/"),
         project = "IYS",
         cruise = paste(project, "GoA2019", sep = ":"), 
         station = paste(cruise, TOW_NUMBER, sep="_Stn"),
         trawl = paste(station, "trawl", sep=":"))

# For OBIS it is useful to parse the eventDate and extract year, month and day from it, for visualizations purposes.
trawl2019$year <- as.numeric(format(as.Date(trawl2019$eventDate), "%Y"))
trawl2019$month <- as.numeric(format(as.Date(trawl2019$eventDate), "%m"))
trawl2019$day <- as.numeric(format(as.Date(trawl2019$eventDate), "%d"))

trawl2019_allCatch <- read_excel(here("Trawl", "2019", "raw_data", 
                                      "2019_GoA_Fish_Trawl_catchdata.xlsx"), sheet = "CATCH_FINAL") %>%
  mutate(project = "IYS",
         cruise = paste(project, "GoA2019", sep = ":"),
         station = paste(cruise, `TOW_NUMBER (number)`, sep = "_Stn"),
         trawl = paste(station, "trawl", sep = ":"))

trawl2019_specimen <- read_excel(here("Trawl", "2019", "raw_data", "2019_GoA_Fish_Specimen_data.xlsx"), 
                                 sheet = "SPECIMEN_FINAL") %>%
  mutate(project = "IYS",
         cruise = paste(project, "GoA2019", sep = ":"),
         station = paste(cruise, TOW_NUMBER, sep = "_Stn"),
         trawl = paste(station, "trawl", sep = ":"),
         sample = paste(trawl, "sample", sep = ":"),
         sample = paste(sample, row_number(), sep = ":"))
```

***

## Event Core 

Next, dataframes are created for each unique `eventID` layer, with unique associated columns. Eventually, the data frames get joined, and the eventCore gets saved locally and in Google Drive. The column `eventRemarks` is not a required column, but includes descriptive information pertaining to any issues that arose during the setting of the trawl net. 

``` {r trawl_event, eval = FALSE}
trawl2019_project <- trawl2019 %>%
  select(eventID = project) %>%
  distinct(eventID) %>%
  mutate(type = "project")

trawl2019_cruise <- trawl2019 %>% 
  select(eventID = cruise,
         parentEventID = project) %>% 
  distinct(eventID, .keep_all = TRUE) %>%
  mutate(type = "cruise")

trawl2019_station <- trawl2019 %>% 
  select(eventID = station,
         parentEventID = cruise) %>% 
  distinct(eventID, .keep_all = TRUE) %>%
  mutate(type = "station")

# The coordinates associated to the trawl need to be presented in a LINESTRING:
trawl2019_coordinates <- trawl2019 %>%
  select(eventID = trawl,
         START_LATITUDE_DD,
         longitude,
         END_LATITUDE_DD,
         END_LONGITUDE_DD) %>%
  mutate(END_LONGITUDE_DD = END_LONGITUDE_DD * -1,
         footprintWKT = paste("LINESTRING (", longitude, START_LATITUDE_DD, ",", 
                              END_LONGITUDE_DD, END_LATITUDE_DD, ")")) 

trawl2019_linestring <- obistools::calculate_centroid(trawl2019_coordinates$footprintWKT)
trawl2019_linestring <- cbind(trawl2019_coordinates, trawl2019_linestring) %>%
  select(eventID, footprintWKT, decimalLatitude, decimalLongitude, coordinateUncertaintyInMeters)

trawl2019_trawl <- trawl2019 %>% 
  select(eventID = trawl,
         parentEventID = station,
         eventDate,
         year,
         month,
         day,
         fieldNotes = PROBLEM_DESCRIPTION) %>%
  mutate(minimumDepthInMeters = 0, # trawl was at the surface
         maximumDepthInMeters = max(trawl2019$MOUTH_OPENING_WIDTH),
         samplingProtocol = "surface trawl", # if possible add DOI to paper here
         locality = case_when(
           trawl2019$EVENT_SUB_TYPE == "Can EEZ" ~ "Canadian EEZ"),
         locationID = case_when(
           trawl2019$EVENT_SUB_TYPE == "Can EEZ" ~ "http://marineregions.org/mrgid/8493")) %>%
  left_join(trawl2019_linestring, by = "eventID") %>% 
  distinct(eventID, .keep_all = TRUE) %>%
    mutate(type = "surface trawl")
  
trawl2019_sample <- trawl2019_specimen %>%
  select(eventID = sample,
         parentEventID = trawl) %>%
  distinct(eventID, .keep_all = TRUE) %>%
  mutate(type = "individual sample")

trawl2019_event <- bind_rows(trawl2019_project, 
                             trawl2019_cruise,
                             trawl2019_station,
                             trawl2019_trawl,
                             trawl2019_sample) 

col_order <- c("eventID", "parentEventID", "eventDate", "year", "month", "day", "decimalLatitude", "decimalLongitude", "footprintWKT", "coordinateUncertaintyInMeters", "minimumDepthInMeters", "maximumDepthInMeters", "samplingProtocol", "locality", "locationID", "type", "fieldNotes")
trawl2019_event <- trawl2019_event[, col_order]

# Flatten event records:
trawl2019_event <- obistools::flatten_event(trawl2019_event)

# Save locally and on Google Drive:
write_csv(trawl2019_event, here("Trawl", "2019", "tidy_data", "trawl2019_event.csv"))
drive_upload(here("Trawl", "2019", "tidy_data", "trawl2019_event.csv"),
             path = "",
             name = "trawl2019_eventCore.csv",
             overwrite = TRUE)
```

## Occurrence

As the trawl data consists of biotic data, an occurrence extension table is created, whereby unique `occurrenceIDs` get linked to the `eventID`. The `occurrenceID` is created from the `eventID`, and any measurements pertaining to the overall catch _or_ specific individual will be associated to that `occurrenceID` in the extended MeasurementOrFact (eMOF) extension.

The Occurrence extension will consist of two tables merged, one from the aggregated (summary) catch data (`catch_final`), and one from the individual length and weight measurements. There will be unique `occurrenceIDs` related to the trawl layer in Event Core. Measurements associated with this `occurrenceID` will be total species' catch biomass and total number caught. The `occurrenceID` for the aggregated catch data will have the format `IYS:GoA2019_Stn1:trawl:occ-1`. An additional layer in the Event Core (sample) was created so that other unique `occurrenceIDs` can be linked to that layer, with `occurrenceIDs` having the format `IYS:GoA2019_Stn1:trawl:sample1:occ-1`. Measurements linked to these `occurrenceIDs` are individual length and weight. For a diagram on the relational database structure, see [here](https://lucid.app/lucidchart/a1d199b4-ef42-45fc-a384-273925fef7d7/edit?page=0_0#). Additionally, a column `occurrenceRemarks` will be included to indicate that that specific `occurrenceID` is a subset of the aggregated catch data. 

For discussion on how `occurrenceIDs` are linked to different layers, see [here](https://github.com/iobis/env-data/issues/10).

A required column for the Occurrence extension is the term, `occurrenceStatus`, which is used to determine presence/absence of data. The package `worrms` is used to generate hierarchical taxonomic information on the species observed. For the `scientificNameID`, we look up the species on the WoRMS website (http://www.marinespecies.org/). 

Finally, we can do a quick check to make sure that all the unique species are included in the trawl_occ dataframe. We need to make sure that the output is TRUE. If confirmed, the Occurrence extension can be saved locally and in the correct Google Drive folder. We also need to make sure that _if_ columns 'SPECIES' or 'SPECIES (Description if needed)' make mention of a lifestage, or sex, we need to include this in the occurrence extension under separate columns. 

The column `species` in DwC terms is referred to as the `specificEpithet`.

``` {r trawl2020_occ, eval = FALSE}
trawl2019_allCatch <- trawl2019_allCatch %>% 
  mutate(scientificname = paste(genus, species, sep = " "))

# The scientificname includes a lot of 'sp.', which often is not read in properly by the worrms package. Therefore, this has to be removed, and included as an identificationQualifier. Once included in the identificationQualifier, remove from scientificname. 
trawl2019_allCatch$identificationQualifier <- ifelse(grepl("\\b sp.\\b|\\b sp\\b", trawl2019_allCatch$scientificname),
                                                     "sp. inc.", NA)
trawl2019_allCatch$scientificname <- gsub("\\b sp.\\b|\\b sp\\b", "", trawl2019_allCatch$scientificname)
trawl2019_allCatch$species <- gsub("sp.|sp", NA, trawl2019_allCatch$species)
         
trawl2019_allCatch_sn <- worrms::wm_records_names(unique(trawl2019_allCatch$scientificname)) %>% bind_rows()
trawl2019_allCatch_taxa <- left_join(trawl2019_allCatch, trawl2019_allCatch_sn, by = "scientificname")

# You'll notice that not all unique_taxa have associated taxonomic hierarchial information. Let's inspect these observations as they might be either misspelled, or a non-Latin name is provided. 
unidentified_species <- trawl2019_allCatch_taxa %>% filter(is.na(AphiaID)) %>% distinct(scientificname)
```

A total of 6 species could not be matched to a record within the WoRMS database. Most often this is due to spelling errors, the inclusion of '.sp', or the lack of supplying a Latin name. I manually check these species in the WoRMS database and confirm with the data provider that the suggested species are the ones meant. These species will then be corrected in the trawl2019_allCatch_taxa data frame. 

- [ ] TO DO: Confirm corrections with Chrys Neville. 

``` {r trawl2020_occ cont, eval = FALSE}
trawl2019_allCatch$scientificname <- gsub("Phacellophora camtshchatica", "Phacellophora camtschatica", 
                                          trawl2019_allCatch$scientificname)
trawl2019_allCatch$scientificname <- gsub("Ctenephores", "Ctenophora", trawl2019_allCatch$scientificname)
trawl2019_allCatch$scientificname <- gsub("Moroteuthis robusta", "Moroteuthis robustus", trawl2019_allCatch$scientificname)
trawl2019_allCatch$scientificname <- gsub("Chrysaora melonaster", "Chrysaora melanaster", trawl2019_allCatch$scientificname)
trawl2019_allCatch$scientificname <- gsub("Paralepididae gen. sp. 1", "Paralepididae", trawl2019_allCatch$scientificname)
trawl2019_allCatch$scientificname <- gsub("NA Siphonophora", "Siphonophora", trawl2019_allCatch$scientificname)
  
# So now we can run worrms::wm_records_names again, and there should be no NA in the AphiaID column.
trawl2019_allCatch_worms <- worrms::wm_records_names(unique(trawl2019_allCatch$scientificname)) %>% bind_rows()
trawl2019_occ <- left_join(trawl2019_allCatch, trawl2019_allCatch_worms, by = "scientificname") %>%
  rename(eventID = trawl,
         specificEpithet = species,
         scientificNameAuthorship = authority,
         taxonomicStatus = status,
         taxonRank = rank,
         scientificName = scientificname,
         scientificNameID = lsid) %>%
  mutate(occurrenceID = paste(eventID, "occ", sep = ":"),
         occurrenceID = paste(occurrenceID, row_number(), sep = "-"),
         occurrenceStatus = "present",
         lifeStage = NA, 
         sex = NA)

# By joining those two dataframes, the column 'genus' is duplicated. Make sure to select the proper genus column, the one generated from the worrms package: 
trawl2019_occ_fnl <- trawl2019_occ %>%
  select(eventID, occurrenceID, occurrenceStatus,
         specificEpithet, scientificName, scientificNameID, identificationQualifier,
         scientificNameAuthorship, taxonomicStatus, taxonRank, kingdom, phylum, 
         class, order, family, genus.y, lifeStage, sex)
```

So now we have created the first Occurrence extension table, which comes from the overall catch data. In the trawl2019_allCatch dataframe, column `Length (N)` indicates how many individuals at each trawl were measured for their lengths. In the next section, we create a second occurrence extension table, with `occurrenceID`s created for each unique individual whose length has been measured. A column will have to be added to indicate that the individual measurements are a subset of the overall catch at that station (`occurrenceRemarks`?). 

``` {r occ_species, eval = FALSE}
trawl2019_catch_ind <- trawl2019_specimen %>%
  rename(eventID = sample) %>%
  mutate(scientificname = paste(genus, species, sep = " "))

trawl2019_catch_ind_sn <- worrms::wm_records_names(unique(trawl2019_allCatch$scientificname)) %>% bind_rows()
trawl2019_catch_ind_taxa <- left_join(trawl2019_catch_ind, trawl2019_catch_ind_sn, by = "scientificname")

# Lets inspect whether there are any unidentified species:
unidentified_species_ind <- trawl2019_catch_ind_taxa %>% filter(is.na(AphiaID)) %>% distinct(scientificname)

trawl2019_catch_ind$identificationQualifier <- ifelse(grepl("\\b sp.\\b|\\b sp. 1\\b", trawl2019_catch_ind$scientificname),
                                                     "sp. inc.", NA)
trawl2019_catch_ind$scientificname <- gsub("\\b sp.\\b|\\b sp. 1\\b", "", trawl2019_catch_ind$scientificname)


trawl2019_catch_ind$scientificname <- gsub("Moroteuthis robusta", "Moroteuthis robustus", trawl2019_catch_ind$scientificname)
trawl2019_catch_ind$scientificname <- gsub("Chrysaora melonaster", "Chrysaora melanaster", trawl2019_catch_ind$scientificname)

# So now we can run worrms::wm_records_names again, and there should be no NA in the AphiaID column.
trawl2019_catch_ind_worms <- worrms::wm_records_names(unique(trawl2019_catch_ind$scientificname)) %>% bind_rows()

trawl2019_catch_ind_occ <- left_join(trawl2019_catch_ind, trawl2019_catch_ind_worms, by = "scientificname") %>%
  rename(scientificNameAuthorship = authority,
         taxonomicStatus = status,
         taxonRank = rank,
         scientificName = scientificname,
         scientificNameID = lsid) %>%
  mutate(occurrenceID = paste(eventID, "occ", sep = ":"),
         occurrenceID = paste(occurrenceID, row_number(), sep = "-"),
         occurrenceStatus = "present",
         occurrenceRemarks = NA)

# By joining those two dataframes, the column 'genus' is duplicated. Make sure to select the proper genus column, the one generated from the worrms package: 
trawl2019_catch_ind_fnl <- trawl2019_catch_ind_occ %>%
  select(eventID, occurrenceID, occurrenceStatus, scientificName, scientificNameID,
         scientificNameAuthorship, taxonomicStatus, taxonRank, kingdom, phylum, 
         class, order, family, genus.y, occurrenceRemarks, identificationQualifier,
         SEX_CODE)

# sex and lifestage is recorded in the SEX_CODE column: 
trawl2019_catch_ind_fnl <- trawl2019_catch_ind_fnl %>%
  mutate(sex = case_when(
    SEX_CODE == "M" ~ "male",
    SEX_CODE == "F" ~ "female",
    SEX_CODE == "N" ~ "unknown"),
         lifeStage = case_when(
    SEX_CODE == "J" ~ "juvenile",
    SEX_CODE == "3" ~ " ")) %>%
  select(-SEX_CODE)
```

- [ ] TO DO: Ask Chrys Neville what SEX_CODE = 3 implies. 

To finalize the Occurrence extension, combine the two occurrence tables. I prefer to re-order it so that it visually makes sense to me. Then, save the combined Occurrence extension locally and in Google Drive. 

``` {r trawl occurrence final, eval = FALSE}
trawl2019_occ_ext <- dplyr::bind_rows(trawl2019_occ_fnl, trawl2019_catch_ind_fnl)

# To re-order the occurrenceID, use following code:
order <- stringr::str_sort(trawl2019_occ_ext$occurrenceID, numeric=TRUE)
trawl2019_occ_ext <- trawl2019_occ_ext[match(order, trawl2019_occ_ext$occurrenceID),]

# Save the occurrence Core locally and in Google Drive:
write_csv(trawl2019_occ_ext, here("Trawl","2019", "tidy_data", "trawl2019_occ_final.csv"))
```

To quality control whether each species in the original specimen file has been accounted for in the final, combined occurrence extension, use the following chunk of code. We also have to check whether the unique species listed in the specimen file match the unique species listed in the overall catch file, where numbers have been aggregated. 

``` {r trawl2020_occ_check}
trawl2019_event_occ <- trawl2019_event %>% filter(!is.na(eventDate))
obistools::plot_map_leaflet(trawl2019_event_occ) # Issue: Station 20 does not have lats and longs associated
```

***

## extended MeasurementOrFact (eMOF) Extension

We create a few measurements or fact tables initially - one for the measurements pertaining to the overall catch (catch_weight and total catch numbers), and one with the measurements pertaining to selected individuals (individual length and weight). These tables will eventually be merged into a single extended measurementOrFact (eMOF) extension. 

Additionally, we need to create eMOF tables with measurements pertaining to the sampling event: tow speed, etc. These come from the `trawl2019` dataset. Variables in the samplingEffort category are `tow_speed`, `tow_duration`, `tow_distance` and `tow_direction`. There is also data collected on bottom depth, every 15 minutes. We save these tables separately, and also combine them to form one final eMOF extension. Through the IPT, multiple eMOF extensions can be uploaded if preferred. 

```{r bottomdepth, eval = FALSE}
trawl2019_bottomdepth <- trawl2019 %>%
  select(eventID = trawl,
         BOTTOM_DEPTH_START_METERS,
         BOTTOM_DEPTH_END_METERS) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = BOTTOM_DEPTH_START_METERS:BOTTOM_DEPTH_END_METERS,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
                                  BOTTOM_DEPTH_START_METERS = "bottom_depth_trawl_start",
                                  BOTTOM_DEPTH_END_METERS = "bottom_depth_trawl_finish"),
         measurementTypeID = case_when(
           measurementType == "bottom_depth_trawl_start" ~ "http://vocab.nerc.ac.uk/collection/P07/current/CFSN0481/",
           measurementType == "bottom_depth_trawl_finish" ~ "http://vocab.nerc.ac.uk/collection/P07/current/CFSN0481/"),
         # Perhaps this is better: http://vocab.nerc.ac.uk/collection/A05/current/EV_BATHY/
         measurementUnit = "meters"),
         measurementUnitID = case_when(
           measurementUnit == "meters" ~ "http://vocab.nerc.ac.uk/collection/P06/current/ULAA/"),
         measurementValueID = NA,
         measurementID = paste(eventID, measurementType, sep = ":")) %>%
  select(eventID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)
# Save the bottom depth extension locally:
write_csv(trawl_bottomdepth, here("Trawl", "tidy_data", "trawl2020_bottomdepth_emof.csv"))
```

Measurements pertaining to the sampling effort: 

```{r samplingEffort, eval = FALSE}
trawl2019_samplingEffort <- trawl2019 %>%
  select(eventID = trawl,
         TOW_DURATION,
         SPEED,
         DISTANCE_TRAVELLED,
         BEARING) %>%
  mutate(TOW_DURATION = round(TOW_DURATION * 60)) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = TOW_DURATION:BEARING,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
         TOW_DURATION = "tow_duration",
         BEARING = "direction_of_set",
         SPEED = "vessel_speed",
         DISTANCE_TRAVELLED = "tow_distance"),
         measurementTypeID = case_when(
           measurementType == "tow_duration" ~ "http://vocab.nerc.ac.uk/collection/P01/current/AZDRZZ01/",
           measurementType == "direction_of_set" ~ "http://vocab.nerc.ac.uk/collection/P07/current/CFSN0481/",
           measurementType == "vessel_speed" ~ "http://vocab.nerc.ac.uk/collection/P01/current/TOWSPEED/",
           measurementType == "tow_distance" ~ "http://vocab.nerc.ac.uk/collection/P01/current/LENTRACK/"),
         measurementUnit = case_when(
           measurementType == "tow_duration" ~ "minutes",
           measurementType == "direction_of_set" ~ "degrees",
           measurementType == "vessel_speed" ~ "knots",
           measurementType == "tow_distance" ~ "nautical miles"),
         measurementUnitID = case_when(
           measurementUnit == "minutes" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UMIN/",
           measurementUnit == "degrees" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UABB/",
           measurementUnit == "knots" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UKNT/",
           measurementUnit == "nautical miles" ~ " "),
         measurementValueID = NA,
         measurementID = paste(eventID, measurementType, sep = ":")) %>%
  select(eventID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)

# Save the extended measurementOrFact extension locally and in Google Drive:
write_csv(trawl2019_samplingEffort, here("Trawl", "tidy_data", "trawl2020_samplingEffort_emof.csv"))
```

And chunk of code to assign sampling instrument measurements/facts:

```{r samplingInstrument, eval = FALSE}
trawl2019_samplingInstrument <- trawl2019 %>%
  select(eventID = trawl,
         GEAR_CODE,
         MOUTH_OPENING_WIDTH,
         MOUTH_OPENING_HEIGHT) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = GEAR_CODE:MOUTH_OPENING_HEIGHT,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
         GEAR_CODE = "sampling_gear",
         MOUTH_OPENING_WIDTH = "net_opening_width",
         MOUTH_OPENING_HEIGHT = "net_opening_height"),
         measurementTypeID = case_when(
           measurementType == "sampling_gear" ~ "http://vocab.nerc.ac.uk/collection/Q01/current/Q0100002/",
           measurementType == "net_opening_width" ~ "http://vocab.nerc.ac.uk/collection/Q01/current/Q0100014/",
           measurementType == "net_opening_height" ~ "http://vocab.nerc.ac.uk/collection/Q01/current/Q0100013/"),
         measurementUnit = case_when(
           measurementType != "sampling_gear" ~ "meters"),
         measurementUnitID = case_when(
           measurementUnit == "meters" ~ "http://vocab.nerc.ac.uk/collection/P06/current/ULAA/"),
         measurementValueID = case_when(
           measurementType == "sampling_gear" ~ " "),
         measurementID = paste(eventID, measurementType, sep = ":")) %>%
  select(eventID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)

# Save the extended measurementOrFact extension locally and in Google Drive:
write_csv(trawl2019_samplingInstrument, here("Trawl", "tidy_data", "trawl2020_samplingInstrument_emof.csv"))
```

And a chunk of code to assign environmental conditions during the trawl:

```{r environmental conditions, eval = FALSE}
trawl2019_conditions <- trawl2019 %>%
  select(eventID = trawl,
         WIND_DIRECTION,
         WIND_SPEED,
         SEA_STATE_ID,
         WATER_TEMPERATURE) %>%
  mutate(SWELL_HEIGHT = ifelse(grepl("swell", trawl2019_conditions$SEA_STATE_ID), trawl2019_conditions$SEA_STATE_ID, NA),
         SWELL_HEIGHT = gsub("swell ", "", trawl2019_conditions$SWELL_HEIGHT),
         SEA_STATE_ID = ifelse(grepl("swell", trawl2019_conditions$SEA_STATE_ID), NA, trawl2019_conditions$SEA_STATE_ID)) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = WIND_DIRECTION:SWELL_HEIGHT,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
                                  SEA_STATE_ID = "wave_height",
                                  WATER_TEMPERATURE = "water_temperature",
                                  WIND_DIRECTION = "wind_direction",
                                  WIND_SPEED = "wind_speed",
                                  SWELL_HEIGHT = "swell_height"),
         measurementTypeID = case_when(
           measurementType == "wave_height" ~ "http://vocab.nerc.ac.uk/collection/P07/current/JNQS0CMX/",
           measurementType == "swell_height" ~ "http://vocab.nerc.ac.uk/collection/P09/current/SWHT/",
           measurementType == "water_temperature" ~ "http://vocab.nerc.ac.uk/collection/P01/current/TEMPPR01/",
           measurementType == "wind_direction" ~ "http://vocab.nerc.ac.uk/collection/B39/current/relwinddir/",
           measurementType == "wind_speed" ~ "http://vocab.nerc.ac.uk/collection/P07/current/CFSN0038/"),
         measurementUnit = case_when(
           measurementType == "wave_height" ~ "meters",
           measurementType == "swell_height" ~ "meters",
           measurementType == "wind_direction" ~ "degrees",
           measurementType == "wind_speed" ~ "knots",
           measurementType == "water_temperature" ~ "degrees celsius"),
         measurementUnitID = case_when(
           measurementUnit == "meters" ~ "http://vocab.nerc.ac.uk/collection/P06/current/ULAA/",
           measurementUnit == "degrees" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UAAA/",
           measurementUnit == "knots" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UKNT/",
           measurementUnit == "degrees celsius" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UPAA/"),
         measurementValueID = NA,
         measurementID = paste(eventID, measurementType, sep = ":")) %>% 
  drop_na(measurementValue) %>%
  select(eventID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)
  

# Save the extended measurementOrFact extension locally and in Google Drive:
write_csv(trawl2019_conditions, here("Trawl", "tidy_data", "trawl2019_conditions_emof.csv"))
```

Next two eMOF extensions are created with measurements pertaining to overall species' catch weight at each trawl, and individual length and weight measurements.

```{r eMoF}
trawl2019_allcatch_measurement <- trawl2019_occ %>%
  select(occurrenceID,
         organismQuantity,
         Mass_net) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = c(organismQuantity, Mass_net),
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
                                  Mass_net = "Total species biomass in net",
                                  organismQuantity = "individualCount"),
         measurementID = case_when(
           measurementType == "individualCount" ~ paste(occurrenceID, "individualCount", sep = "-"),
           measurementType == "Total species biomass in net" ~ paste(occurrenceID, "biomass", sep = "-")),
         measurementTypeID = case_when(
           measurementType == "Total species biomass in net" ~ "http://vocab.nerc.ac.uk/collection/S06/current/S0600088/",
           measurementType == "individualCount" ~ "http://vocab.nerc.ac.uk/collection/P01/current/OCOUNT01/"),
         measurementUnit = case_when(
           measurementType == "Total species biomass in net" ~ "kilogram",
           measurementType == "individualCount" ~ "individuals"),
         measurementUnitID = case_when(
           measurementUnit == "kilogram" ~ "http://vocab.nerc.ac.uk/collection/P06/current/KGXX/",
           measurementUnit == "individuals" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UUUU/"),
         measurementValueID = NA) %>%
  select(occurrenceID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)
```

Next, create the measurementOrFact extension with length and weight measurements assigned to the _individual_ specimens. Some changes here are made in the Processed Data: The `SEX_CODE` covers both sex (male, female, unknown) and lifestage (J - which is juvenile? or should it be i for immature?)

- [ ] TO DO: Ask Chrys Neville which species were measured what way (total, fork, mantle length). 
- [ ] TO DO: Ask Chrys what `SEX_CODE` = 3 represents.
- [ ] TO DO: Ask Chrys what `SEX_CODE` = 'j' represents - juvenile? Or should it be i for immature (see metadata sheet)?

``` {r}
trawl2019_ind_emof <- trawl2019_catch_ind_occ %>%
  select(occurrenceID,
         LENGTH, `STANDARD LENGTH`, SEX_CODE, `WEIGHT - TOTAL`) %>%
  mutate(lifeStage = ifelse(grepl("3|J", trawl2019_ind_emof$SEX_CODE), trawl2019_ind_emof$SEX_CODE, NA),
         SEX_CODE = gsub("3|J", NA, trawl2019_ind_emof$SEX_CODE)) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = LENGTH:`WEIGHT - TOTAL`,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementType = recode(measurementType,
                                  LENGTH = "species length",
                                  `STANDARD LENGTH` = "Standard length",
                                  SEX_CODE = "sex",
                                  `WEIGHT - TOTAL` = "total weight"),
         measurementID = case_when(
           measurementType == "species length" ~ paste(occurrenceID, "species_length", sep = "-"),
           measurementType == "Standard length" ~ paste(occurrenceID, "standard_length", sep = "-"),
           measurementType == "sex" ~ paste(occurrenceID, "sex", sep = "-"),
           measurementType == "lifeStage" ~ paste(occurrenceID, "lifeStage", sep = "-"),
           measurementType == "total weight" ~ paste(occurrenceID, "total_weight", sep = "-")),
         measurementTypeID = case_when(
           measurementType == "species length" ~ "http://vocab.nerc.ac.uk/collection/P01/current/OBSINDLX/",
           measurementType == "Standard length" ~ "http://vocab.nerc.ac.uk/collection/P01/current/SL01XX01/",
           measurementType == "sex" ~ " ",
           measurementType == "lifeStage" ~ " ",
           measurementType == "total weight" ~ "http://vocab.nerc.ac.uk/collection/S06/current/S0600088/"),
         measurementUnit = case_when(
           measurementType == "species length" ~ "millimeters",
           measurementType == "Standard length" ~ "millimeters",
           measurementType == "total weight" ~ "grams"),
         measurementUnitID = case_when(
           measurementUnit == "millimeters" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UXMM/",
           measurementUnit == "centimeters" ~ "http://vocab.nerc.ac.uk/collection/P06/current/ULCM/",
           measurementUnit == "grams" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UGRM/"),
         measurementValueID = case_when(
           measurementValue == "F" ~ " ",
           measurementValue == "M" ~ " ",
           measurementValue == "N" ~ " ",
           measurementValue == "J" ~ " ",
           measurementValue == "3" ~ " ")) %>%
  mutate(measurementValueID = NA) %>%
  drop_na(measurementValue) %>%
  select(occurrenceID, measurementID, measurementType, measurementTypeID, measurementValue, measurementValueID,
         measurementUnit, measurementUnitID)
```

Combine the two extended measurementOrFact extensions into a single eMOF extension, and save locally and in GoogleDrive:

``` {r}
trawl2020_eMOF <- dplyr::bind_rows(trawl2020_allcatch_emof, trawl2020_ind_emof)
# To combine all the extensions, use: 
# trawl2020_all_emofs <- dplyr::bind_rows(trawl_bottomdepth, samplingEffort, samplingInstrument, trawl_conditions, trawl2020_allcatch_emof, trawl2020_ind_emof)
write_csv(trawl2020_eMOF, here("Trawl", "tidy_data", "trawl_eMOF.csv"))
drive_upload(here("Trawl", "tidy_data", "trawl_emof.csv"),
             path = "https://drive.google.com/drive/folders/1ZbvwjD8odJr2SUKq6Z8zAwpLd6QeOOYC",
             name = "trawl_eMoF.csv",
             overwrite = TRUE)
```