---
title: "CTD Data wrangle"
author: "Tim van der Stap & Julian Gan"
date: "5/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(readxl)
library(here)
library(parsedate)
library(googledrive)
```

Download the raw data from GoogleDrive, and read from local folder:

``` {r file_download, eval = FALSE}
drive_download("https://drive.google.com/open?id=1-onECiDW02DJRIX6g4dXywtzLEqCGyce", path = here::here("CTD", "raw_data", "Data_IYS_conbined_Final.xlsx"))

sheet1 <- read_excel(here("CTD", "raw_data", "Data_IYS_conbined_Final.xlsx"), sheet = "Sheet1")
sheet2 <- read_excel(here("CTD", "raw_data", "Data_IYS_conbined_Final.xlsx"), sheet = "Sheet2")
```

## EventCore 

The reported date / time has to get formatted to UTC in the ISO8601 extended format. Additionally, to wrangle the data into a tidy format that follows the Darwin Core Archive (DwC-A) standards and hierarchical structure, it's important to determine how many layers of `eventID` (and, consequently, `parentEventID`) the data has. In the CTD dataset, the data has four layers: "cruise", "station", "cast", and the depth at which the measurements were taken ("ndepth"). Additional columns for each of these layers are created. 

``` {r time_formatting, eval = FALSE}
sheet2 <- read_excel(here("CTD", "raw_data", "Data_IYS_conbined_Final.xlsx"), sheet = "Sheet2") %>%
  mutate(eventDate_text = str_c(YEAR, MONTH, DAY, `TIME(ship time)`, `MIN(ship time)`, sep="-"),
         eventDate = ymd_hm(eventDate_text, tz = "Asia/Kamchatka")) %>%
  mutate(eventDate = format_iso_8601(as.POSIXct(eventDate,
                                                format = "%Y-%m-%d %H:%M%:S",
                                                tz="Asia/Kamchatka")),
         eventDate = str_replace(eventDate, "\\+00:00", "Z")
         ) %>%
  mutate(cruise = "GoA2019",
         station = paste(cruise, `NO.Trawl`, sep="_Stn"),
         cast = paste(station, "cast1", sep = ":"),
         ndepth = paste(cast, DEPTH, sep=":ctd:"))
```

The longitude was recorded as positive values, however, after QC and cross-reference to other data recorded during the same cruise, it is clear that the values should be negative.

``` {r QC_Longitude, eval = FALSE}
sheet2$LON <- -sheet2$LON
```

For each of the layers of `eventID`, a data frame is created with unique eventIDs, associated parentEventIDs, and associated unique characteristics. In our data, we join the date to the station level, as all station-specific data was collected on a day. 

Finally, these tables get joined, and re-ordered based on `eventID`, to create the Event Core. This table gets saved locally and in the correct GoogleDrive folder.

``` {r ctd_event, eval = FALSE}
ctd_cruise <- sheet2 %>% 
  select(eventID = cruise) %>%
  distinct(eventID) %>%
  mutate(type = "cruise")

# Join date to station
ctd_station <- sheet2 %>%
  select(eventID = station,
         parentEventID = cruise,
         eventDate,
         decimalLatitude = LAT,
         decimalLongitude = LON) %>% 
  distinct(eventID, .keep_all = TRUE) %>% 
  mutate(type = "station")

ctd_cast <- sheet2 %>% 
  select(eventID = cast,
         parentEventID = station) %>% 
  distinct(eventID, .keep_all = TRUE) %>% 
  mutate(type = "cast")

ctd_ndepth <- sheet2 %>% 
  select(eventID = ndepth,
         parentEventID = cast,
         minimumDepthInMetres = DEPTH,
         maximumDepthInMetres = DEPTH) %>% 
  distinct(eventID, .keep_all = TRUE) %>% 
  mutate(type = "sample")

ctd_event <- bind_rows(ctd_cruise, ctd_station, ctd_cast, ctd_ndepth) %>% 
  select(eventID, parentEventID:maximumDepthInMetres, type)  

# If you wish to re-order the Event Core, use the following code:
# order <- stringr::str_sort(ctd_event$eventID, numeric=TRUE)
# ctd_event <- ctd_event[match(order, ctd_event$eventID),]

# Save locally and in GoogleDrive
write_csv(ctd_event, here("CTD", "tidy_data", "CTD_event.csv"))
drive_upload(here("CTD", "tidy_data", "CTD_event.csv"),
             path = "https://drive.google.com/drive/u/0/folders/1-XXOPhMN4-BmhI3owM2hvaMKXEy6hEYL",
             name = "CTD_event.csv",
             overwrite = TRUE)
```

## MeasurementOrFact 

After creating the Event Core, we create an extended MeasurementOrFact (eMoF) table. An occurrence table is not created as the datasets only contain abiotic data, therefore, measurements are linked to eventIDs. DwC-A standards work with "long" format, rather than the wide format, so we need to transpose the data. 

Our first step however is to add the column of bottom depth at each station to the extension measurementOrFact table. Please note that the `measurementTypeID` for seafloor depth is defined as "Bridge log sheets" until there is more information about how bottom depth was determined/measured (e.g., by bathymetric sensors, charts, etc.). 

``` {r seafloor, eval = FALSE}
ctd_botdepth <- sheet2 %>% 
  select(eventID = cast, `Bot. Depth`) %>%
  distinct(eventID, .keep_all = TRUE) %>%
  pivot_longer(cols = `Bot. Depth`,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(measurementValue = as.character(measurementValue),
         measurementID = paste(eventID, "depth", sep = ":"),
         measurementType = recode(measurementType,
                                  `Bot. Depth` = "seafloor depth"),
         measurementTypeID = "http://vocab.nerc.ac.uk/collection/C00/current/BRIDGE/",
         measurementUnit = "m",
         measurementUnitID = "http://vocab.nerc.ac.uk/collection/P06/current/ULAA/" ) %>%
  select(measurementID, eventID, measurementType, measurementTypeID, measurementValue,
         measurementUnit, measurementUnitID)
```

Next, we select the appropriate columns and use the pivot_longer function to transpose the data. As EC25 is measured in microSiemens/cm, which is not (yet) in NERCs controlled vocabulary, we convert this column to milliSiemens/cm. 

``` {r measurement data_wrangle, eval = FALSE}
ctd_measurement <- sheet2 %>% 
  select(eventID = ndepth, 
         `TEM_S`:`BOD5[ml/l]`) %>%
  mutate(`EC25 [mS/cm]_R` = as.numeric(`EC25 [uS/cm]_R`) / 1000) %>%
  mutate_all(as.character)
ctd_measurement <- ctd_measurement[, -which(names(ctd_measurement) == "EC25 [uS/cm]_R")] %>%
  pivot_longer(`TEM_S`:`EC25 [mS/cm]_R`,
               names_to = "measurementType",
               values_to = "measurementValue")
```

__Note: in the previous chunk we got our data from sheet2 - this is data collected at continuous depths [every meter]. To create a measurement table for sheet1 (discrete depth measurements), simply replace sheet2 with sheet1.__

Assign NERC controlled vocabulary to the variables and the measurement units. Finally, join the two dataframes and save the eMoF table locally and on GoogleDrive in the correct folder.  

``` {r CTD, eval = FALSE}
ctd_measurement <- ctd_measurement %>% # Should "NO.Trawl" here be "NO.(ST)"?
  mutate(measurementID = case_when(measurementType == "TEM_S" ~ paste(eventID,"TEM_S",sep="_"),
                                   measurementType == "TEM_R" ~ paste(eventID,"TEM_R",sep="_"),
                                   measurementType == "SAL_S" ~ paste(eventID,"SAL_S",sep="_"),
                                   measurementType == "SAL_R" ~ paste(eventID,"SAL_R",sep="_"),
                                   measurementType == "Cond. [mS/cm]_R" ~ paste(eventID,"Cond. [mS/cm]_R", sep="_"),
                                   measurementType == "Density [kg/m^3]_R" ~ paste(eventID,"Density", sep="_"),
                                   measurementType == "SigmaT [ ]_R" ~ paste(eventID,"Sigma",sep="_"),
                                   measurementType == "Chl-Flu. [ppb]_R" ~ paste(eventID,"Chl-Flu", sep="_"),
                                   measurementType == "Chl-a [ug/l]_R" ~ paste(eventID,"Chl-a", sep="_"),
                                   measurementType == "Turb-M [FTU]_R" ~ paste(eventID,"Turbidity",sep="_"),
                                   measurementType == "DO [mg/l]_R" ~ paste(eventID,"DO",sep="_"),
                                   measurementType == "Batt. [V]_R" ~ paste(eventID,"Batt", sep="_"),
                                   measurementType == "pH" ~ paste(eventID,"pH",sep="_"),
                                   measurementType == "O2 [ml/l]" ~ paste(eventID,"O2 [ml/l]",sep="_"),
                                   measurementType == "BOD5[ml/l]" ~ paste(eventID,"BOD5[ml/l]",sep="_"),
                                   measurementType == "DO [%]_R" ~ paste(eventID,"BOD5[ml/l]",sep="_"),
                                   measurementType == "EC25 [mS/cm]_R" ~ paste(eventID,"EC25 [mS/cm]_R",sep="_")),
         
         measurementType = recode(measurementType, 
                                  `TEM_S` = "Sea Water Temperature [Sea-Bird]",
                                  `TEM_R` = "Sea Water Temperature [Rinko]",
                                  `SAL_R` = "Sea Water Practical Salinity [Rinko]",
                                  `SAL_S` = "Sea Water Practical Salinity [Sea-Bird]",
                                  `pH` = "pH",
                                  `O2 [ml/l]` = "Dissolved oxygen [Lab]",
                                  `Cond. [mS/cm]_R` = "Sea Water Electrical Conductivity",
                                  `Density [kg/m^3]_R` = "Sea Water Density",
                                  `SigmaT [ ]_R` = "Sea Water Sigma T",
                                  `Chl-Flu. [ppb]_R` = "Chlorophyll - Flu",
                                  `Chl-a [ug/l]_R` = "Chlorophyll-a",
                                  `Turb-M [FTU]_R` = "Turbidity",
                                  `DO [mg/l]_R` = "Dissolved oxygen",
                                  `Batt. [V]_R` = "Battery voltage",
                                  `BOD5[ml/l]` = "BOD in 5m depth [Lab]",
                                  `DO [%]_R` = "Percent oxygen saturation",
                                  `EC25 [mS/cm]_R` = "EC25"),
         
         measurementTypeID = recode(measurementType, #URIs from NERC website
                                    "Sea Water Temperature [Sea-Bird]" = "http://vocab.nerc.ac.uk/collection/P02/current/TEMP/2/",
                                    "Sea Water Temperature [Rinko]" = "http://vocab.nerc.ac.uk/collection/P02/current/TEMP/2/",
                                    "Sea Water Practical Salinity [Rinko]" = "http://vocab.nerc.ac.uk/collection/P02/current/PSAL/",
                                    "Sea Water Practical Salinity [Sea-Bird]" = "http://vocab.nerc.ac.uk/collection/P02/current/PSAL/",
                                    "Sea Water Electrical Conductivity" = "http://vocab.nerc.ac.uk/collection/P02/current/CNDC/",
                                    "Sea Water Density" = "http://vocab.nerc.ac.uk/collection/P04/current/G990/",
                                    "Sea Water Sigma T" = "http://vocab.nerc.ac.uk/collection/OG1/current/SIGTHETA/",     
                                    "Chlorophyll - Flu" = "http://vocab.nerc.ac.uk/collection/P07/current/CFSN0705/",
                                    "Chlorophyll-a" = "http://vocab.nerc.ac.uk/collection/P07/current/CFSN0705/",
                                    "Turbidity" = "http://vocab.nerc.ac.uk/collection/P25/current/TURB/",
                                    "Dissolved oxygen" = "http://vocab.nerc.ac.uk/collection/P35/current/EPC00002/",
                                    "Battery voltage" = "http://vocab.nerc.ac.uk/collection/S06/current/S0600163/",
                                    "pH" = "http://vocab.nerc.ac.uk/collection/P01/current/PHXXZZXX/",
                                    "Dissolved oxygen [Lab]" = "http://vocab.nerc.ac.uk/collection/P01/current/DOXYZZ01/",
                                    "BOD in 5m depth [Lab]" = "http://vocab.nerc.ac.uk/collection/P01/current/BODZZZZZ/1/",
                                    "Percent oxygen saturation" = "http://vocab.nerc.ac.uk/collection/P02/current/DOXY/",
                                    "EC25" = "http://vocab.nerc.ac.uk/collection/P02/current/CNDC/"),
         
         measurementUnit = case_when(measurementType == "Sea Water Temperature [Sea-Bird]" ~ "deg C",
                                     measurementType == "Sea Water Temperature [Rinko]" ~ "deg C",
                                     measurementType == "Sea Water Practical Salinity [Rinko]" ~ " ",
                                     measurementType == "Sea Water Practical Salinity [Sea-Bird]" ~ " ",
                                     measurementType == "pH" ~ " ",
                                     measurementType == "Sea Water Electrical Conductivity" ~ "mS/cm",
                                     measurementType == "Sea Water Density" ~ "kg m^-3 ",
                                     measurementType == "Sea Water Sigma T" ~ "kg m^-3",
                                     measurementType == "Chlorophyll - Flu" ~ "ppb",
                                     measurementType == "Chlorophyll-a" ~ "ug/L",
                                     # Alternatively, chlorophyll-a concentration can be reported in mg.m-3. Conversion = 1:1.
                                     measurementType == "Turbidity" ~ "FTU",
                                     measurementType == "Battery voltage" ~ "V",
                                     measurementType == "Dissolved oxygen" ~ "mg/L",
                                     measurementType == "Dissolved oxygen [Lab]" ~ "ml/L ",
                                     measurementType == "BOD in 5m depth [Lab]" ~ "ml/L",
                                     measurementType == "Percent oxygen saturation" ~ "percent saturation",
                                     measurementType == "EC25" ~ "mS/cm"),
         
         measurementUnitID = case_when(measurementType == "Sea Water Temperature [Sea-Bird]" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UPAA/",
                                       measurementType == "Sea Water Temperature [Rinko]" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UPAA/",
                                       measurementType == "Sea Water Practical Salinity [Rinko]" ~ "http://vocab.nerc.ac.uk/collection/P06/current/PPTR/",
                                       measurementType == "Sea Water Practical Salinity [Sea-Bird]" ~ "http://vocab.nerc.ac.uk/collection/P06/current/PPTR/",
                                       measurementType == "Sea Water Electrical Conductivity" ~ "http://vocab.nerc.ac.uk/collection/P06/current/MSCM/",
                                       measurementType == "Sea Water Density" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UKMC/",
                                       measurementType == "Sea Water Sigma T" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UKMC/",
                                       measurementType == "Chlorophyll - Flu" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UPPB/",
                                       measurementType == "Chlorophyll-a" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UGPL/",
                                       measurementType == "Turbidity" ~ "http://vocab.nerc.ac.uk/collection/P06/current/USTU/",
                                       measurementType == "Dissolved oxygen" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UMGL/",
                                       measurementType == "Battery voltage" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UVLT/",
                                       measurementType == "pH" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UUPH/",
                                       measurementType == "Dissolved oxygen [Lab]" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UMLL/",
                                       measurementType == "BOD in 5m depth [Lab]" ~ "http://vocab.nerc.ac.uk/collection/P06/current/UMLL/",
                                       measurementType == "Percent oxygen saturation" ~ "http://vocab.nerc.ac.uk/collection/P01/current/OXYSZZ01/",
                                       measurementType == "EC25" ~ "http://vocab.nerc.ac.uk/collection/P06/current/MSCM/")
  ) %>%
  select(measurementID, measurementType, measurementTypeID,
         measurementValue, measurementUnit, measurementUnitID)

# Join the dataframes: 
ctd_measurementOrFact <- bind_rows(ctd_botdepth, ctd_measurement) %>%
  select(measurementID, eventID, measurementType, measurementTypeID,
         measurementValue, measurementUnit, measurementUnitID)

# Save locally and on GoogleDrive: 
write_csv(ctd_measurementOrFact, here("CTD", "tidy_data", "CTD_measurement.csv"))
drive_upload(here("CTD", "tidy_data", "CTD_measurement.csv"),
             path = "https://drive.google.com/drive/u/0/folders/1-XXOPhMN4-BmhI3owM2hvaMKXEy6hEYL",
             name = "CTD_measurement.csv",
             overwrite = TRUE)
```