---
title: "Transforming Nutrient Data to OBIS-DwC"
author: "Julian Gan"
date: "06/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(tidyverse)
library(lubridate)
library(readxl)
library(parsedate)
library(googledrive)
library(here)
```

## Getting started

The following line only needs to be run once to download the data from Google Drive to your computer's hard drive

```{r drive_download}
drive_download("https://drive.google.com/open?id=1MQysK1BEp9xcZH64ilq63Sx3ri5SJYCx", path = here("nutrients", "raw_data", "IYS_GoA_2019_Hydro&Chemistry_20190326.xlsx"))
```

The Excel file contains two tabs: `Hydro_GoA` contains temperature and salinity measurements collected by the Sea-bird CTD at discrete depths. The `Chemistry-GoA` file contains the actual nutrient data (Si, DIP, DIN, NO<sub>2</sub>, and NO<sub>3</sub>), along with oxygen measurements and pH. This script focuses on only transforming the nutrient data to OBIS-DwC. Dates, however, will be pulled from the `Hydro_GoA` tab because there is no temporal data in `Chemistry_GoA`.

Next, we are going to create new columns that will eventually become our `eventID`s. We are going to have 5 different types of `eventID`s, in order of descending hierarchy: "cruise", "station", "cast", "ndepth" (depth at which the sample was taken with a niskin), and "sample" (one sample per niskin collection). The eventIDs follow a sematic model that describes the hierarchical nature of each event type. Each sample is given an `eventID` because the "niskin" `eventID` should be interoperable with other datasets.

```{r excel_wrangle}
# Create a primary key in the Hydro_GoA sheet so that the date column can be joined to Chemistry_GoA
hydro <-
  read_excel(
    here(
      "nutrients",
      "raw_data",
      "IYS_GoA_2019_Hydro&Chemistry_20190326.xlsx"
    ),
    sheet = "Hydro_GoA"
  ) %>%
  mutate(station = paste("GoA2019", Station, sep = "_Stn"))

# Create columns for eventIDs in Chemistry_GoA
chemistry <-
  read_excel(
    here(
      "nutrients",
      "raw_data",
      "IYS_GoA_2019_Hydro&Chemistry_20190326.xlsx"
    ),
    sheet = "Chemistry_GoA"
  ) %>%
  mutate(
    cruise = "GoA2019",
    station = paste(cruise, Station, sep = "_Stn"),
    cast = paste(station, "cast1", sep = ":"),
    ndepth = paste(cast, `Depth [m]`, sep = ":niskin:"),
    sample = paste(ndepth, "N1", sep = ":")
  )
```


## Create the Event core

```{r event}
nut_cruise <- chemistry %>%
  select(eventID = cruise) %>%
  distinct(eventID) %>%
  mutate(type = "cruise")

# eventDate is an attribute associated with the "station" record type
nut_station <- chemistry %>%
  left_join(select(hydro,
                   station,
                   date)) %>%
  select(
    eventID = station,
    parentEventID = cruise,
    eventDate = date,
    decimalLatitude = Latitude,
    decimalLongitude = Longitude
  ) %>%
  distinct(eventID, .keep_all = TRUE) %>%
  mutate(decimalLongitude = decimalLongitude-360,
         type = "station")

nut_cast <- chemistry %>%
  select(eventID = cast,
         parentEventID = station) %>%
  distinct(eventID, .keep_all = TRUE) %>%
  mutate(type = "cast")

nut_ndepth <- chemistry %>%
  select(
    eventID = ndepth,
    parentEventID = cast,
    minimumDepthInMetres = `Depth [m]`,
    maximumDepthInMetres = `Depth [m]`
  ) %>%
  distinct(eventID, .keep_all = TRUE) %>%
  mutate(type = "sample")

nut_sample <- chemistry %>%
  select(eventID = sample,
         parentEventID = ndepth) %>%
  distinct(eventID, .keep_all = TRUE) %>%
  mutate(type = "subsample")

nut_event <-
  bind_rows(nut_cruise, nut_station, nut_cast, nut_ndepth, nut_sample) %>%
  select(eventID, parentEventID:maximumDepthInMetres, type)

# The .gitignore file has been configured to ignore all files in the tidy_data folder so no data gets pushed to the Git repository.
write_csv(nut_event, here("nutrients", "tidy_data", "nut_event.csv"))

# drive_upload can't overwrite files based on name alone, so if you're re-running this script after creating the first file, make sure to delete the older version from Google Drive.
drive_upload(here("nutrients", "tidy_data", "nut_event.csv"),
             path = "https://drive.google.com/drive/u/0/folders/1M0JRuzW_rxwz9jZPVV-oDv5AARUS2K2w",
             name = "nutrient_event.csv")
```


## Create the ExtendedMeasurementorFact extension 

The eMoF extension links measurement records with records in the Event core table. Since this dataset is comprised only of environmental abiotic data, there is no `occurrenceID` column and records are linked by `eventID` instead. The `measurementTypeID` for seafloor depth is defined as "Bridge log sheets" until there is more information about how bottom depth was determined/measured (e.g., by bathymetric sensors, charts, etc.). The measurement unit for all variables was specified in the raw data file as "mkM/l"; this was determined to be a Russian equivalent of "umol/L" ([Vyunova et al .2016](https://www.researchgate.net/publication/311473228_Synacton_and_individual_activity_of_synthetic_and_natural_corticotropins)). Where no specific NERC vocabulary exists for a particular measurement type, I used a broader or related term, as annotated in the code below.

```{r measurements}
# Seafloor (bottom) depth is a measurement of the niskin cast.
nut_botdepth <- chemistry %>%
  select(eventID = cast, `Bot. Depth [m]`) %>%
  distinct(eventID, .keep_all = TRUE) %>%
  pivot_longer(cols = `Bot. Depth [m]`,
               names_to = "measurementType",
               values_to = "measurementValue") %>%
  mutate(
    measurementValue = as.numeric((measurementValue)),
    measurementID = paste(eventID, "depth", sep = ":"),
    measurementType = recode(measurementType,
                             "Bot. Depth [m]" = "seafloor depth"),
    measurementTypeID = "http://vocab.nerc.ac.uk/collection/C00/current/BRIDGE/",
    measurementUnit = "m",
    measurementUnitID = "http://vocab.nerc.ac.uk/collection/P06/current/ULAA/"
  ) %>%
  select(
    measurementID,
    eventID,
    measurementType,
    measurementTypeID,
    measurementValue,
    measurementUnit,
    measurementUnitID
  )


# These are measurements taken from the niskin water samples.
nut_measurement <- chemistry %>%
  select(eventID = sample,
         `Si [mkM/l]`:`NO3  [mkM/l]`) %>%
  pivot_longer(
    cols = `Si [mkM/l]`:`NO3  [mkM/l]`,
    names_to = "measurementType",
    values_to = "measurementValue"
  ) %>%
  mutate(
    measurementType = recode(
      measurementType,
      "Si [mkM/l]" = "silicate",
      "DIP [mkM/l]" = "dissolved inorganic P",
      "DIN [mkM/l]" = "dissolved inorganic N",
      "NO2  [mkM/l]" = "nitrogen dioxiode",
      "NO3  [mkM/l]" = "nitrate",
    ),
    measurementTypeID = case_when(
      measurementType == "silicate" ~ "http://vocab.nerc.ac.uk/collection/P01/current/SLCAZZXX/",
      measurementType == "dissolved inorganic P" ~ "http://vocab.nerc.ac.uk/collection/P01/current/PHOSXXDZ/",
      # Cannot find NERC vocab for DIP; only PO4
      measurementType == "dissolved inorganic N" ~ "http://vocab.nerc.ac.uk/collection/P01/current/DINXZZZZ/",
      measurementType == "nitrogen dioxiode" ~ "http://vocab.nerc.ac.uk/collection/S27/current/CS003170/",
      # No NERC vocab for concentration of NO2 per unit volume of the water body; only of atmosphere
      measurementType == "nitrate" ~ "http://vocab.nerc.ac.uk/collection/P01/current/NTRAXXDZ/"
    ),
    measurementUnit = "umol/L",
    measurementUnitID = "http://vocab.nerc.ac.uk/collection/P06/current/UPOX/",
    measurementID = case_when(
      measurementType == "silicate" ~ paste(eventID, "si", sep = ":"),
      measurementType == "dissolved inorganic P" ~ paste(eventID, "DIP", sep =
                                                           ":"),
      measurementType == "dissolved inorganic N" ~ paste(eventID, "DIN", sep =
                                                           ":"),
      measurementType == "nitrogen dioxiode" ~ paste(eventID, "NO2", sep =
                                                       ":"),
      measurementType == "nitrate" ~ paste(eventID, "NO3", sep =
                                             ":"),
    )
  )

nut_measurementOrFact <-
  bind_rows(nut_botdepth, nut_measurement) %>%
  select(
    measurementID,
    eventID,
    measurementType,
    measurementTypeID,
    measurementValue,
    measurementUnit,
    measurementUnitID
  )

write_csv(nut_measurementOrFact,
          here("nutrients", "tidy_data", "nut_measurement.csv"))
drive_upload(
  here("nutrients", "tidy_data", "nut_measurement.csv"),
  path = "https://drive.google.com/drive/u/0/folders/1M0JRuzW_rxwz9jZPVV-oDv5AARUS2K2w",
  name = "nutrient_eMoF.csv"
)
```

