---
title: "Untitled"
author: "Brett"
date: "24/02/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# IYS High Seas Data Management Recommendations

## Introduction

Objective: Recommend options for the collection, integration
and storage of high seas multi-disciplinary data.

## Key technologies

## Review of current data requirements

```{r}
# Consider using ODBC R package to run sql queries to airtable

# Otherwise there's an R pckage called airtabler https://itsalocke.com/blog/how-to-use-an-r-interface-with-airtable-api/
```


### Physical

### Chenmical

### Biological and Ecosystems

## Assessment of options for data collection, integration and storage in 2021

## The attributes of the envisioned system of data collection include:

* Data collection will be standardized across multiple vessels.
* Data collection at sea will be electronic. Samples inventories should function as per Laboratory Information Systems (LIMS) with the capacity to input results post-cruise.
* There will be capacity to download in realtime during the cruise associated climate
and ocean data that may inform the program at sea.
* Data will be integrated into a shared information system and made freely available

## Resources 

[NPAFC’s Bering-Aleutian Salmon International Survey-II (BASIS-II)](https://npafc.org/basis-phase-ii/)

